### 2016-12-10 article.tex
First version

%  LaTeX support: latex@mdpi.com 
%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

% You need to save the "mdpi.cls" and "mdpi.bst" files into the same folder as this template file.

%=================================================================
\documentclass[journal,article,submit,moreauthors,pdftex,10pt,a4paper]{mdpi}

%
%--------------------
% Class Options:
%--------------------
% journal
%----------
% Choose between the following MDPI journals:
% actuators, admsci, aerospace, agriculture, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, atmosphere, atoms, axioms, batteries, behavsci, beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci, buildings, carbon, cancers, catalysts, cells, challenges, chemosensors, children, chromatography, climate, coatings, computation, computers, condensedmatter, cosmetics, cryptography, crystals, data, dentistry, designs, diagnostics, diseases, diversity, econometrics, economies, education, electronics, energies, entropy, environments, epigenomes, fermentation, fibers, fishes, fluids, foods, forests, futureinternet, galaxies, games, gels, genealogy, genes, geosciences, geriatrics, healthcare, horticulturae, humanities, hydrology, informatics, information, infrastructures, inorganics, insects, instruments, ijerph, ijfs, ijms, ijgi, ijtpp, inventions, jcdd, jcm, jdb, jfb, jfmk, jimaging, jof, jintelligence, jlpea, jmse, jpm, jrfm, jsan, land, languages, laws, life, literature, lubricants, machines, magnetochemistry, marinedrugs, materials, mathematics, mca, mti, medsci, medicines, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, molbank, molecules, mps, nanomaterials, ncrna, neonatalscreening, nutrients, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, plants, polymers, processes, proteomes, publications, recycling, religions, remotesensing, resources, risks, robotics, safety, scipharm, sensors, separations, sexes, sinusitis, socsci, societies, soils, sports, standards, sustainability, symmetry, systems, technologies, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vetsci, viruses, vision, water
%---------
% article
%---------
% The default type of manuscript is article, but can be replaced by: 
% addendum, article, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, correction, conferenceproceedings, conferencereport, expressionofconcern, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimage, letter, newbookreceived, opinion, obituary, projectreport, reply, retraction, review, perspective, preprints, shortnote, supfile, technicalnote, viewpoint
% supfile = supplementary materials
%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g. the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figure are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother 
\articlenumber{x}
\doinum{10.3390/------}
\pubvolume{xx}
\pubyear{2016}
\copyrightyear{2016}
\externaleditor{Academic Editor: name}
\history{Received: date; Accepted: date; Published: date}

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Remark, Definition
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Title}

% If this is an expanded version of a conference paper, please cite it here: enter the full citation of your conference paper, and add $^\dagger$ in the end of the title of this article.
%\conference{}

% Authors, for the paper (add full first names)
\Author{Firstname Lastname $^{1,\dagger,\ddagger}$, Firstname Lastname $^{1,\ddagger}$ and Firstname Lastname $^{2,}$*}
% Authors, for metadata in PDF
\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}

% Contact information of the corresponding author
\corres{Correspondence: e-mail@e-mail.com; Tel.: +x-xxx-xxx-xxxx}

% Current address and/or shared authorship
\firstnote{Current address: Affiliation 3} 
\secondnote{These authors contributed equally to this work.}

% Simple summary
%\simplesumm{}

% Abstract (Do not use inserted blank lines, i.e. \\) 
\abstract{A single paragraph of about 200 words maximum. For research articles, abstracts should give a pertinent overview of the work. We strongly encourage authors to use the following style of structured abstracts, but without headings: 1) Background: Place the question addressed in a broad context and highlight the purpose of the study; 2) Methods: Describe briefly the main methods or treatments applied; 3) Results: Summarize the article's main findings; and 4) Conclusion: Indicate the main conclusions or interpretations. The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}

% Keywords
\keyword{keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.)}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:

%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% For Conference Proceedings Papers: add the conference title here
%\conferencetitle{}

%\setcounter{secnumdepth}{4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Only for the journal Gels: Please place the Experimental Section after the Conclusions

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{section}{-1} %% Remove this when starting to work on the template.
\section{How to Use this Template}
The template details the sections that can be used in a manuscript. Note that the order of article sections may differ from the requirements of the journal (e.g. for the positioning of the Materials and Methods section). Please check the instructions for authors page of the journal to verify the correct order. For any questions, please contact the editorial office of the journal or support@mdpi.com. For LaTeX related questions please contact Janine Daum at latex-support@mdpi.com.

\section{Introduction}

The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper \cite{ref-journal}. And now citing a book reference \cite{ref-book}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Subsection}

\subsubsection{Subsubsection}

Bulleted lists look like this:
\begin{itemize}[leftmargin=*,labelsep=4mm]
\item	First bullet
\item	Second bullet
\item	Third bullet
\end{itemize}

Numbered lists can be added as follows:
\begin{enumerate}[leftmargin=*,labelsep=3mm]
\item	First item
\item	Second item
\item	Third item
\end{enumerate}

The text continues here.

\subsection{Figures, Tables and Schemes}

All figures and tables should be cited in the main text as Figure 1, Table 1, etc.

\begin{figure}[H]
\centering
\includegraphics[width=3cm]{logo-mdpi}
\caption{This is a figure, Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.}
\end{figure}

\begin{table}[H]
\caption{This is a table caption. Tables should be placed in the main text near to the first time they are cited.}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{ccc}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
entry 1		& data			& data\\
entry 2		& data			& data\\
\bottomrule
\end{tabular}
\end{table}

\subsection{Formatting of Mathematical Components}

This is an example of an equation:

\begin{equation}
\mathbb{S}
\end{equation}

%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph. 
Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
%% Example of a theorem:
\begin{Theorem}
Example text of a theorem.
\end{Theorem}

The text continues here. Proofs must be formatted as follows:

%% Example of a proof:
\begin{proof}[Proof of Theorem 1]
Text of the proof. Note that the phrase `of Theorem 1' is optional if it is clear which theorem is being referred to.
\end{proof}
The text continues here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

Authors should discuss the results and how they can be interpreted in perspective of previous studies and of the working hypotheses. The findings and their implications should be discussed in the broadest context possible. Future research directions may also be highlighted.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}

Materials and Methods should be described with sufficient details to allow others to replicate and build on published results. Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers. Please disclose at the submission stage any restrictions on the availability of materials or information. New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.

Research manuscripts reporting large datasets that are deposited in a publicly available database should specify where the data have been deposited and provide the relevant accession numbers. If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\supplementary{The following are available online at www.mdpi.com/link, Figure S1: title, Table S1: title, Video S1: title.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{All sources of funding of the study should be disclosed. Please clearly indicate grants that you have received in support of your research work. Clearly state if you received funds for covering the costs to publish in open access.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``X.X. and Y.Y. conceived and designed the experiments; X.X. performed the experiments; X.X. and Y.Y. analyzed the data; W.W. contributed reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be limited to those who have contributed substantially to the work reported.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictofinterests{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results. Any role of the funding sponsors in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section. If there is no role, please state ``The founding sponsors had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, and in the decision to publish the results''.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\abbreviations{The following abbreviations are used in this manuscript:\\

\noindent 
\begin{tabular}{@{}ll}
MDPI & Multidisciplinary Digital Publishing Institute\\
DOAJ & Directory of open access journals\\
TLA & Three letter acronym\\
LD & linear dichroism
\end{tabular}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixsections{multiple} %Leave argument "multiple" if there are multiple sections. Then a counter is printed ("Appendix A?). If there is only one appendix section then change the argument to ?one? and no counter is printed (?Appendix?).
\appendix
\section{}
The appendix is an optional section that can contain details and data supplemental to the main text. For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data. Mathemtaical proofs of results not central to the paper can be added as an appendix.

\section{}
All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here. 
\bibliographystyle{mdpi}

%=====================================
% References, variant A: internal bibliography
%=====================================
\renewcommand\bibname{References}
\begin{thebibliography}{999}
% Reference 1
\bibitem{ref-journal}
Lastname, F.; Author, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142-149.
% Reference 2
\bibitem{ref-book}
Lastname, F.F.; Author, T. The title of the cited contribution. In {\em The Book Title}; Editor, F., Meditor, A., Eds.; Publishing House: City, Country, 2007; pp. 32-58.
\end{thebibliography}

%=====================================
% References, variant B: external bibliography
%=====================================
%\bibliography{your_external_BibTeX_file}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\sampleavailability{Samples of the compounds ...... are available from the authors.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

### 2016-12-13 article.tex
New article draft

@@
 %=================================================================
-\documentclass[journal,article,submit,moreauthors,pdftex,10pt,a4paper]{mdpi}
+\documentclass[entropy,article,submit,moreauthors,pdftex,10pt,a4paper]{mdpi}
 
@@
 % Choose between the following MDPI journals:
-% actuators, admsci, aerospace, agriculture, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, atmosphere, atoms, axioms, batteries, behavsci, beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci, buildings, carbon, cancers, catalysts, cells, challenges, chemosensors, children, chromatography, climate, coatings, computation, computers, condensedmatter, cosmetics, cryptography, crystals, data, dentistry, designs, diagnostics, diseases, diversity, econometrics, economies, education, electronics, energies, entropy, environments, epigenomes, fermentation, fibers, fishes, fluids, foods, forests, futureinternet, galaxies, games, gels, genealogy, genes, geosciences, geriatrics, healthcare, horticulturae, humanities, hydrology, informatics, information, infrastructures, inorganics, insects, instruments, ijerph, ijfs, ijms, ijgi, ijtpp, inventions, jcdd, jcm, jdb, jfb, jfmk, jimaging, jof, jintelligence, jlpea, jmse, jpm, jrfm, jsan, land, languages, laws, life, literature, lubricants, machines, magnetochemistry, marinedrugs, materials, mathematics, mca, mti, medsci, medicines, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, molbank, molecules, mps, nanomaterials, ncrna, neonatalscreening, nutrients, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, plants, polymers, processes, proteomes, publications, recycling, religions, remotesensing, resources, risks, robotics, safety, scipharm, sensors, separations, sexes, sinusitis, socsci, societies, soils, sports, standards, sustainability, symmetry, systems, technologies, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vetsci, viruses, vision, water
+% actuators, admsci, aerospace, agriculture, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, atmosphere, atoms, axioms, batteries, behavsci, beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci, buildings, carbon, cancers, catalysts, cells, challenges, chemosensors, children, chromatography, climate, coatings, computation, computers, condensedmatter, cosmetics, cryptography, crystals, data, dentistry, designs, diagnostics, diseases, diversity, econometrics, economies, education, electronics, energies, entropy, environments, epigenomes, fermentation, fibers, fishes, fluids, foods, forests, futureinternet, galaxies, games, gels, genealogy, genes, geosciences, geriatrics, healthcare, horticulturae, humanities, hydrology, informatics, information, infrastructures, inorganics, insects, instruments, ijerph, ijfs, ijms, ijgi, ijtpp, inventions, jcdd, jcm, jdb, jfb, jfmk, jimaging, jof, jintelligence, jlpea, jmse, jpm, jrfm, jsan, land, languages, laws, life, literature, lubricants, machines, magnetochemistry, marinedrugs, materials, mathematics, mca, mti, medsci, medicines, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, molbank, molecules, mps, nanomaterials, ncrna, neonatalscreening, nutrients, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, plants, polymers, processes, proteomes, publications, recycling, religions, remotesensing, resources, risks, robotics, safety, sensors, separations, sexes, sinusitis, socsci, societies, soils, sports, standards, sustainability, symmetry, systems, technologies, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vetsci, viruses, water
 %---------
@@
 % The default type of manuscript is article, but can be replaced by: 
-% addendum, article, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, correction, conferenceproceedings, conferencereport, expressionofconcern, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimage, letter, newbookreceived, opinion, obituary, projectreport, reply, retraction, review, perspective, preprints, shortnote, supfile, technicalnote, viewpoint
+% addendum, article, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, correction, conferenceproceedings, conferencereport, expressionofconcern, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimage, letter, newbookreceived, opinion, obituary, projectreport, reply, retraction, review, preprints, shortnote, supfile, technicalnote
 % supfile = supplementary materials
@@
 % Full title of the paper (Capitalized)
-\Title{Title}
-
-% If this is an expanded version of a conference paper, please cite it here: enter the full citation of your conference paper, and add $^\dagger$ in the end of the title of this article.
-%\conference{}
+\Title{EEG characterization and classification based on the histograms of visual gradients of plots}
 
 % Authors, for the paper (add full first names)
-\Author{Firstname Lastname $^{1,\dagger,\ddagger}$, Firstname Lastname $^{1,\ddagger}$ and Firstname Lastname $^{2,}$*}
+\Author{Rodrigo Ramele $^{1,\dagger,\ddagger}$, Ana Julia Villar $^{1,\ddagger}$ and Juan Miguel Santos  $^{1,\ddagger}$*}
 % Authors, for metadata in PDF
-\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}
+\AuthorNames{Rodrigo Ramele, Ana Julia Villar and Juan Miguel Santos}
 
 % Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
-\address{%
-$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
-$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}
+\address[1]{%
+$^{1}$ \quad Computer Engineering Department, Instituto Tecnológico de Buenos Aires ITBA; info@itba.edu.ar}
 
 % Contact information of the corresponding author
-\corres{Correspondence: e-mail@e-mail.com; Tel.: +x-xxx-xxx-xxxx}
+\corres{Correspondence: rramele@itba.edu.ar; Tel.: +54-9-11-4193-9382}
 
 % Current address and/or shared authorship
-\firstnote{Current address: Affiliation 3} 
+\firstnote{Current address: Lavarden 315, Postal Code C1437FBH, City of Buenos Aires, Argentina} 
 \secondnote{These authors contributed equally to this work.}
@@
 We strongly encourage authors to use the following style of structured abstracts, but without headings: 1) Background: Place the question addressed in a broad context and highlight the purpose of the study; 2) Methods: Describe briefly the main methods or treatments applied; 3) Results: Summarize the article's main findings; and 4) Conclusion: Indicate the main conclusions or interpretations.
-The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}
+The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.This is great.
+We had tremendous results.
+Our results are so great that nobody will believe that we actually managed to have them !}
 
 % Keywords
-\keyword{keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.)}
+\keyword{EEG; BCI; P300; ALS; classification; hog; sift}
 
@@
 %\JEL{}
+%\AMS{}
+
+% If this is an expanded version of a conference paper, please cite it here: enter the full citation of your conference paper, and add $^\S$ in the end of the title of this article.
+%\conference{}
 
@@
 
+One of the most shoking reasons is that we can use the capacity of ML techniques to automatically explore the signal, by using these new featrues.
+
 \section{Introduction}
@@
 And now citing a book reference \cite{ref-book}.
+
+In this work EEG signals from ALS patients are offline processed with this radically different and horizontal approach.  This is the continuation of the work presented here and here.
+
+The same methodological algorithm can be used to process rhythmic EEG markers or paradigms and at the same time to analyze spatiotemporal phenomena like ERP, particularly P300.
+
+This method is based on a morphological analysis of the shape of the EEG signal.
+
+P300 single trial is a golden grail of Brain computer interfaces.  It is studied and analyzed in this way, here in this other way, in this other way is here.
 
@@
 \centering
-\includegraphics[width=3cm]{logo-mdpi}
+\includegraphics[width=9cm]{subjectaveraged.eps}
 \caption{This is a figure, Schemes follow the same formatting.
@@
 
-\noindent 
-\begin{tabular}{@{}ll}
-MDPI & Multidisciplinary Digital Publishing Institute\\
-DOAJ & Directory of open access journals\\
-TLA & Three letter acronym\\
-LD & linear dichroism
-\end{tabular}}
+\noindent MDPI: Multidisciplinary Digital Publishing Institute\\
+DOAJ: Directory of open access journals\\
+TLA: Three letter acronym\\
+LD: linear dichroism}
 

### 2016-12-18 article.tex
The paper is materializing.....

@@
 \address[1]{%
-$^{1}$ \quad Computer Engineering Department, Instituto Tecnológico de Buenos Aires ITBA; info@itba.edu.ar}
+$^{1}$ \quad Computer Engineering Department, Instituto Tecnológico de Buenos Aires (ITBA); info@itba.edu.ar}
 
@@
 % Current address and/or shared authorship
-\firstnote{Current address: Lavarden 315, Postal Code C1437FBH, City of Buenos Aires, Argentina} 
+\firstnote{Current address: C1437FBH Lavarden 315, Ciudad Autónoma de Buenos Aires, Argentina} 
 \secondnote{These authors contributed equally to this work.}
@@
 \setcounter{section}{-1} %% Remove this when starting to work on the template.
-\section{How to Use this Template}
-The template details the sections that can be used in a manuscript.
-Note that the order of article sections may differ from the requirements of the journal (e.g.
-for the positioning of the Materials and Methods section).
-Please check the instructions for authors page of the journal to verify the correct order.
-For any questions, please contact the editorial office of the journal or support@mdpi.com.
-For LaTeX related questions please contact Janine Daum at latex-support@mdpi.com.
-
-One of the most shoking reasons is that we can use the capacity of ML techniques to automatically explore the signal, by using these new featrues.
 
@@
 
-The introduction should briefly place the study in a broad context and highlight why it is important.
-It should define the purpose of the work and its significance.
-The current state of the research field should be reviewed carefully and key publications cited.
-Please highlight controversial and diverging hypotheses when necessary.
-Finally, briefly mention the main aim of the work and highlight the principal conclusions.
-As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research.
-Citing a journal paper \cite{ref-journal}.
-And now citing a book reference \cite{ref-book}.
+\begin{enumerate}[leftmargin=*,labelsep=3mm]
+\item	Why
+\item	Purpose and significance
+\item	Current state of the research field and key publications cited
+\item   Controversial hypothesis
+\item   Main aim and principal conclusions.
+\end{enumerate}
 
-In this work EEG signals from ALS patients are offline processed with this radically different and horizontal approach.  This is the continuation of the work presented here and here.
+Cybathlon
 
-The same methodological algorithm can be used to process rhythmic EEG markers or paradigms and at the same time to analyze spatiotemporal phenomena like ERP, particularly P300.
+x) El paper tiene un foco general en EEG y rapidamente se centra en BCI como una prueba del método.
+x) La relevancia de estudiar métodos de análisis de EEG para pacientes con ALS.
+x) Ofrece una Receta de cómo implementar un esquema simple de detección de P300 y ofrece una comparativa contra el paper de las italianas.
+x) La transversalidad del método de poder detectar eventos rítmicos asi como transientes como p300.
+x) La posibilidad de utilizarlo para reconocer, fuertemente un patron particular y con la flexibilidad de detectar variantes basado en las poblaciones de descriptores (esta es mi hipótesis de porque funciona, porque yo vi que el método se hace pelota con pequeñas variantes de la señal).
+x) La importancia del averaging de EEG y sus desbalanceos, puntualizando que en el paper original
+x) Que la caída en la clasificación no cae significativamente aún bajando la cantidad de repeticiones requeridas para el averaging.
+x) ¿ Por qué da mejor en PO8 que en Cz y Pz que son los canales P300 ?  (como future work).
+
+Although recently impressive advances in neuroimagining techniques has pushed EEG, Electroencephalography technology a little bit out of the way in clinical context, where it has always being used since its discovery by Berger.
+
+Their versatility, ease of use, temporal resolution, ease and development of fabrication, and the explosion as a consumer devices, is pushing EEG devices to become the de facto non invasive method to access and harness brain information.
+
+Key contribution to this development has been the field of Brain Computer Interfaces, which is the development of a new channel of communication particularly aimed to person with some degree of neuro degenerative disease.
+
+One very remarkable aspect of this communication channel is the ability to transmit some general cognitive state, like alertness, drowsiness, boredom, and so on, which can very helpful particularly in rehabilitation procedures.
+\citep{Guger2009a}.
+
+One of the most shocking reasons is that we can use the capacity of ML techniques to automatically explore the signal, by using these new features.
+
+Additionally, they have a high variability between different subjects and even between different moments for the same subject.
+This inherent complexity comes from  is a real challenge when it is required to extract information from raw EEG signals.
+
+Signal is complex and it is difficult to find those values which are not variable.
+
+Due to this inner complexity, it is often necessary to implement many distinct and specialized algorithmic methods, to filter the signal, classify it, and try to determine some meaning out of it.
+
+In this work EEG signals from ALS patients are offline processed with this radically different and horizontal approach.
+This is the continuation of the work presented here and here \citep{Ramele2016}, which was previously applied to rythmic patterns.
+
+The same methodological algorithm can be used to process rhythmic EEG markers or paradigms and at the same time to analyze spatiotemporal phenomena like ERP, particularly P300 \citep{Ramele2016}.
 
@@
 
+Successful applications of BCI has been applied to rehabilitation \citep{Jure2016}.
+
+This paper reports on a study done to address
+the hypothesis that communication task complexity
+(copy-spelling vs. novel text generation of a picture
+description) is correlated with BCI performance accuracy
+and P300 characteristics.
+
+BCI accuracy
+was calculated as the percentage of attempted selections
+that matched the intended selection.
+
+\section{Materials and Methods}
+
+Besides new novel applications of BCI, the core value of the entire discipline is to provide communication assistance to patients.
+It is understood that people suffering from ALS, even though the disease do not harm the CNS, their brain signals and suffer slow and deterioriating changes that lead to more complex patterns to be detected [cite]
+
+Although they are used (PAPER DE P300 donde dice que se usa en los laboratorios) they still lack the bla bla bla CITE LOTTE BOOK \citep{Huggins2016}
+
+We use the dataset 008-2014 published on the BNCI-Horizon website.
+This consortium of companies and universities were developed under the umbrella of the EU to define, provide guidance for BCI development \citep{Riccio2013}.
+
+\subsection{Experimental protocol}
+
+The protocol is explained [HERE] but can be summarized as follows:  8 subjects affected by different levels of ALS disease, were used to generate all the data.
+The P300 task was meant to spell 7 words of 5 letter each.
+The first 3 runs were used for training and the remaining 4 for testing with visual feedback.
+A trial, for this experiment, was defined as each attempt to select a letter from the speller, and it was composed of 10 repetions of flashes of 6 rows and 6 columns of the matrix, yielding 120 repetitions.
+Flashing of row/columns was performed for 0.125 s, following by a resting period of the same size.
+
+This dataset was recorded from a healthy subject during a no-feedback session.
+She/he sat in a normal chair with relaxed arms resting on the table and fingers in the standard typing position at the computer keyboard.
+The task was to press with the index and little fingers the corresponding keys in a self-chosen order and timing 'self-paced key typing'.
+The experiment consisted of 3 sessions of 6 minutes each.
+All sessions were conducted on the same day with some minutes break in-between.
+Typing was done at an average speed of 1 key per second.
+
+\subsection{Morphological Signal Analysis}
+
+Due to the inner complexity of EEG signals, several approaches and methods has been applied to decode them.  Of particular interests are morphological signal analysis 
+Paper del chabon del ITBA que detalla bien los diferentes metodos,
+
+paper de los mexicanos \citep{Alvarado-Gonzalez2016}.
+
+Morphological or
+
+template matching
+
+spike
+sharp wave
+spike wave
+polyspike 
+slow wave complex
+
+shape domain
+
+\subsubsection{P300}
+
+Uhhhh I can talk a lot about P300 \citep{Knuth2006}.
+
 P300 single trial is a golden grail of Brain computer interfaces.  It is studied and analyzed in this way, here in this other way, in this other way is here.
+
+\subsection{Algorithm}
+
+\subsubsection{Preprocessing}
+
+\subsubsection{Signal to Image Transformation}
+
+There are many ways an EEG matrix can be transformed into an image.
+
+What we do is take the signal and plot it.  Thats all folks.
+
+\subsubsection{Signal Averaging}
+
+The methods for averaging.
+
+\subsubsection{Classification}
+
+Finally the classification is carried out by using the NBNN algorithm which can be described as follows.
+
+BCI Simulation vs k-fold cross validation
 
@@
 
-This section may be divided by subheadings.
-It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
+Results are shown in the following chart.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=18cm]{subjectaverageds4.eps}
+\caption{This is a figure, Schemes follow the same formatting.
+If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel.
+(\textbf{b}) Description of what is contained in the second panel.
+Figures should be placed in the main text near to the first time they are cited.
+A caption on a single line should be centered.}
+\end{figure}
+
+Contrary to the literature (\citep{Huggins2016,Jure2016}) the best performing channels were not Cz and Pz and instead Pz.
+
+GRAND AVERAGES
+
+it is quite likely that the data will be unbalanced (that is, an unequal number of trials in each condition)
+
+\begin{table}[H]
+\caption{Results obtained for Accuracy levels by k-fold cross validation.}
+\centering
+%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
+\begin{tabular}{cccc}
+\toprule
+\textbf{Cz}	& \textbf{Cz}	& \textbf{Best Channel}	& \textbf{Performance}\\
+\midrule
+0.845000     & 0.794286 & 2 & 0.794286 $\pm$ 0.011952 \\
+0.863000     & 0.811429 & 7 & 0.934286 $\pm$ 0.011200 \\
+0.872000     & 0.780000 & 2 & 0.780000 $\pm$ 0.031843 \\
+0.859000     & 0.675714 & 8 & 0.904286 $\pm$ 0.013232 \\
+0.862000     & 0.801429 & 7 & 0.932857 $\pm$ 0.008400 \\
+0.886000     & 0.964286 & 2 & 0.964286 $\pm$ 0.010435 \\
+0.886000     & 0.784286 & 7 & 0.930000 $\pm$ 0.009744 \\
+0.923000     & 0.908571 & 7 & 0.997143 $\pm$ 0.003733 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+The ITR, or BTR, in the case of reactive BCIs (CITE WOLPAW) strongly depends on the amount of averaging required to transmit a valid and robust selection.
+If the amount of repetitions is reduced for each subject, performing a BCI Simulation can be shown on Fig.
+\ref{fig:singletrial}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=18cm]{singletriality.eps}
+\caption{This is a figure, Schemes follow the same formatting.
+If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel.
+(\textbf{b}) Description of what is contained in the second panel.
+Figures should be placed in the main text near to the first time they are cited.
+A caption on a single line should be centered.}
+\label{fig:singletrial}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
+
+This method, different from other methods which is based on the nonlinearity of the gradient of histograms which can be used to detect
+
+is also based on how the image look like.
+
+Results of one against the other
+
+SNR of p300 and how to detect it
+
+Check if you can use this to detect any kind of transient signbal
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusion}
+
+With this result we showed a method to process EEG signals where the main characteristic can be both, rhytmic in nature as in motor imagery, and also transient and in time space, like the P300 event related potential.
+
+If we avoid removing artefacts achieve a higher value of classification.
+
+Si el desempeño da parecido hay que escribirlo ya que con menos valores se permite un ITR mayor.
+
+We also achieved single trial classification for just one subject.
 
@@
 \end{figure}
-
-\begin{table}[H]
-\caption{This is a table caption. Tables should be placed in the main text near to the first time they are cited.}
-\centering
-%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
-\begin{tabular}{ccc}
-\toprule
-\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
-\midrule
-entry 1		& data			& data\\
-entry 2		& data			& data\\
-\bottomrule
-\end{tabular}
-\end{table}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\acknowledgments{All sources of funding of the study should be disclosed.
-Please clearly indicate grants that you have received in support of your research work.
-Clearly state if you received funds for covering the costs to publish in open access.}
+\acknowledgments{This project was supported by the ITBACyT-15 funding program issued by ITBA University.}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
-The following statements should be used ``X.X.
-and Y.Y.
-conceived and designed the experiments; X.X.
-performed the experiments; X.X.
-and Y.Y.
-analyzed the data; W.W.
-contributed reagents/materials/analysis tools; Y.Y.
-wrote the paper.'' Authorship must be limited to those who have contributed substantially to the work reported.}
+\authorcontributions{Juan Miguel Santos organized the project and conceived the idea. All the authors equally worked in the completion of this project.}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\conflictofinterests{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results.
-Any role of the funding sponsors in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section.
-If there is no role, please state ``The founding sponsors had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, and in the decision to publish the results''.}
+\conflictofinterests{The authors declare no conflict of interest.}
 
@@
 %=====================================
-\renewcommand\bibname{References}
-\begin{thebibliography}{999}
+%\renewcommand\bibname{References}
+%\begin{thebibliography}{999}
 % Reference 1
-\bibitem{ref-journal}
-Lastname, F.; Author, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142-149.
+%\bibitem{ref-journal}
+%Lastname, F.; Author, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142-149.
 % Reference 2
-\bibitem{ref-book}
-Lastname, F.F.; Author, T. The title of the cited contribution. In {\em The Book Title}; Editor, F., Meditor, A., Eds.; Publishing House: City, Country, 2007; pp. 32-58.
-\end{thebibliography}
+%\bibitem{ref-book}
+%Lastname, F.F.; Author, T. The title of the cited contribution. In {\em The Book Title}; Editor, F., Meditor, A., Eds.; Publishing House: City, Country, 2007; pp. 32-58.
+%\end{thebibliography}
 
@@
 %=====================================
-%\bibliography{your_external_BibTeX_file}
+\bibliography{article}
 

### 2016-12-18 article.tex
General layout of the paper

@@
 
-Cybathlon
-
-x) El paper tiene un foco general en EEG y rapidamente se centra en BCI como una prueba del método.
-x) La relevancia de estudiar métodos de análisis de EEG para pacientes con ALS.
-x) Ofrece una Receta de cómo implementar un esquema simple de detección de P300 y ofrece una comparativa contra el paper de las italianas.
-x) La transversalidad del método de poder detectar eventos rítmicos asi como transientes como p300.
-x) La posibilidad de utilizarlo para reconocer, fuertemente un patron particular y con la flexibilidad de detectar variantes basado en las poblaciones de descriptores (esta es mi hipótesis de porque funciona, porque yo vi que el método se hace pelota con pequeñas variantes de la señal).
-x) La importancia del averaging de EEG y sus desbalanceos, puntualizando que en el paper original
-x) Que la caída en la clasificación no cae significativamente aún bajando la cantidad de repeticiones requeridas para el averaging.
-x) ¿ Por qué da mejor en PO8 que en Cz y Pz que son los canales P300 ?  (como future work).
+\begin{enumerate}[leftmargin=*,labelsep=3mm]
+\item El paper tiene un foco general en EEG y rapidamente se centra en BCI como una prueba del método.
+\item La relevancia de estudiar métodos de análisis de EEG para pacientes con ALS.
+\item Ofrece una Receta de cómo implementar un esquema simple de detección de P300 y ofrece una comparativa contra el paper de las italianas.
+\item La transversalidad del método de poder detectar eventos rítmicos asi como transientes como p300.
+\item La posibilidad de utilizarlo para reconocer, fuertemente un patron particular y con la flexibilidad de detectar variantes basado en las poblaciones de descriptores (esta es mi hipótesis de porque funciona, porque yo vi que el método se hace pelota con pequeñas variantes de la señal).
+\item La importancia del averaging de EEG y sus desbalanceos, puntualizando que en el paper original
+\item Que la caída en la clasificación no cae significativamente aún bajando la cantidad de repeticiones requeridas para el averaging.
+\item ¿ Por qué da mejor en PO8 que en Cz y Pz que son los canales P300 ?  (como future work).
+\end{enumerate}
 
@@
 
-Their versatility, ease of use, temporal resolution, ease and development of fabrication, and the explosion as a consumer devices, is pushing EEG devices to become the de facto non invasive method to access and harness brain information.
+Their versatility, ease of use, temporal resolution, ease and development of fabrication, and the explosion as a consumer devices, is pushing EEG devices to become the de-facto non invasive method to access and harness brain information.
 
@@
 
-One very remarkable aspect of this communication channel is the ability to transmit some general cognitive state, like alertness, drowsiness, boredom, and so on, which can very helpful particularly in rehabilitation procedures.
-\citep{Guger2009a}.
+One very remarkable aspect of this communication channel is the ability to volitionally transmit information from the CNS, Central Nervous System, to a computer device and from there use that information to drive wheelchair or as input in a speller application \citep{Guger2009a}, or a rehabilitation procedure \citep{Jure2016} in connection to FES.
 
-One of the most shocking reasons is that we can use the capacity of ML techniques to automatically explore the signal, by using these new features.
+EEG signals are remarkably complex and has been characterized as a non-stationary stochastic process with even sometimes their baseline activity is near random (CITE TEODIANO BASTOS).
 
 Additionally, they have a high variability between different subjects and even between different moments for the same subject.
-This inherent complexity comes from  is a real challenge when it is required to extract information from raw EEG signals.
+This inherent complexity is a real challenge when it is required to extract information from raw EEG signals.
 
-Signal is complex and it is difficult to find those values which are not variable.
+Hence, inevitably, it is often necessary to implement many distinct and specialized algorithmic methods, to filter the signal, enhance its SNR, and try to determine some meaning out of it.
 
-Due to this inner complexity, it is often necessary to implement many distinct and specialized algorithmic methods, to filter the signal, classify it, and try to determine some meaning out of it.
-
-In this work EEG signals from ALS patients are offline processed with this radically different and horizontal approach.
-This is the continuation of the work presented here and here \citep{Ramele2016}, which was previously applied to rythmic patterns.
+In this work a new method to characterize EEG signals is presented, expanded and detailed.
+Its validity is shown by processing offline data from ALS patients.
+This is the continuation of the work previously presented \citep{Ramele2016}, which was applied mainly to rythmic patterns.
 
@@
 This method is based on a morphological analysis of the shape of the EEG signal.
-
-Successful applications of BCI has been applied to rehabilitation \citep{Jure2016}.
 
@@
 and P300 characteristics.
-
-BCI accuracy
-was calculated as the percentage of attempted selections
-that matched the intended selection.
 
@@
 
-The protocol is explained [HERE] but can be summarized as follows:  8 subjects affected by different levels of ALS disease, were used to generate all the data.
+The protocol is explained \citep{Riccio2013} but can be summarized as follows:  8 subjects affected by different levels of ALS disease, were used to generate all the data.
 The P300 task was meant to spell 7 words of 5 letter each.
 The first 3 runs were used for training and the remaining 4 for testing with visual feedback.
-A trial, for this experiment, was defined as each attempt to select a letter from the speller, and it was composed of 10 repetions of flashes of 6 rows and 6 columns of the matrix, yielding 120 repetitions.
+A trial, for this experiment, was defined as each attempt to select a letter from the speller, and it was composed of 10 repetitions of flashes of 6 rows and 6 columns of the matrix, yielding 120 repetitions.
 Flashing of row/columns was performed for 0.125 s, following by a resting period of the same size.
+After 120 repetions and intertrial pause is performed to start with the following letter.
 
-This dataset was recorded from a healthy subject during a no-feedback session.
-She/he sat in a normal chair with relaxed arms resting on the table and fingers in the standard typing position at the computer keyboard.
-The task was to press with the index and little fingers the corresponding keys in a self-chosen order and timing 'self-paced key typing'.
-The experiment consisted of 3 sessions of 6 minutes each.
-All sessions were conducted on the same day with some minutes break in-between.
-Typing was done at an average speed of 1 key per second.
+All sessions were conducted on the same day with some minutes break in-between. Typing was done at an average speed of 1 key per second.
 
@@
 
-Due to the inner complexity of EEG signals, several approaches and methods has been applied to decode them.  Of particular interests are morphological signal analysis 
-Paper del chabon del ITBA que detalla bien los diferentes metodos,
-
-paper de los mexicanos \citep{Alvarado-Gonzalez2016}.
+Due to the inner complexity of EEG signals, several approaches and methods have been applied to decode them.  Of particular interests are morphological signal analysis 
+CITE (Paper del chabon del ITBA que detalla bien los diferentes metodos) \citep{Alvarado-Gonzalez2016}.
 
@@
 
-Results are shown in the following chart.
+In Fig.
+\ref{fig:subjectaveraged} the grand average (point-to-point) of for all the subjects is shown.
+The P300 characteristic curve can be seen particularly in subjects 6, and 7 and in a lesser extend in the remaining subjects.
+In order to obtain a valid binary classification on averaged signals, particular care was observed to avoid unbalanced samples, because that may introduce a bias in the classification procedure (signal variance is proportional to the root of the number of samples and the procedure would be discriminating signals with different variances).it is quite likely that the data will be unbalanced (that is, an unequal number of trials in each condition)
 
@@
 A caption on a single line should be centered.}
+\label{fig:subjectaveraged}
 \end{figure}
 
-Contrary to the literature (\citep{Huggins2016,Jure2016}) the best performing channels were not Cz and Pz and instead Pz.
-
-GRAND AVERAGES
-
-it is quite likely that the data will be unbalanced (that is, an unequal number of trials in each condition)
+Results are shown in the following chart where BCI accuracy
+was calculated as the percentage of successful predictions
+that matched the proposed selection in a copy-spelling approach (CITE TAYLOR AND  FRANCOS PAPER).
 
 \begin{table}[H]
-\caption{Results obtained for Accuracy levels by k-fold cross validation.}
+\caption{Results obtained for Accuracy levels by k-fold cross validation.
+The values reported for Cz are reproduced here for comparison.
+Cz values of BCI accuracies can be seen as long as the accuracy levels obtained for the best performing channels, for each subject.}
 \centering
@@
 \end{table}
+
+Contrary to the literature (\citep{Huggins2016,Jure2016}) the best performing channels were not Cz and Pz but instead the occipital channels PO1 and P02 (5 against 3).
 
@@
 
-Results of one against the other
-
 SNR of p300 and how to detect it
 
-Check if you can use this to detect any kind of transient signbal
+Check if you can use this to detect any kind of transient signal.
+
+Compare if it is possible with the descriptors from one subject, discriminate the others.
+
+Channal identification based on the metric distance between the bags
 

### 2016-12-19 article.tex
Shaping the skeleton of the paper

@@
 % Keywords
-\keyword{EEG; BCI; P300; ALS; classification; hog; sift}
+\keyword{electroencephalography (EEG); BCI; P300; ALS; classification; HOG; SIFT}
 
@@
 
-Although recently impressive advances in neuroimagining techniques has pushed EEG, Electroencephalography technology a little bit out of the way in clinical context, where it has always being used since its discovery by Berger.
+Although recently impressive advances in neuroimagining (particularly radio-nuclear and radiological scanning methods) \citep{Schomer2010} techniques has pushed the traditional EEG, Electroencephalography technology a little bit out of the way in clinical context, the advent and advances in digitalized counterparts has pushed for a revamping of the entire discipline.
 
-Their versatility, ease of use, temporal resolution, ease and development of fabrication, and the explosion as a consumer devices, is pushing EEG devices to become the de-facto non invasive method to access and harness brain information.
+Their versatility, ease of use, temporal resolution, ease of development and fabrication, and the explosion as consumer devices, is pushing EEG devices to become the de-facto non invasive method to access and harness brain information.
 
@@
 
-One very remarkable aspect of this communication channel is the ability to volitionally transmit information from the CNS, Central Nervous System, to a computer device and from there use that information to drive wheelchair or as input in a speller application \citep{Guger2009a}, or a rehabilitation procedure \citep{Jure2016} in connection to FES.
+One very remarkable aspect of this communication channel is the ability to volitionally transmit information from the CNS, Central Nervous System, to a computer device and from there use that information to drive wheelchair or as input in a speller application \citep{Guger2009a}, or a rehabilitation procedure \citep{Jure2016}.
+The golden goal of BCI is to implement a new complete and alternative pathway to restore motion \citep{WolpawJonathanR2012}.
 
-EEG signals are remarkably complex and has been characterized as a non-stationary stochastic process with even sometimes their baseline activity is near random (CITE TEODIANO BASTOS).
+EEG signals are remarkably complex and has been characterized as a non-stationary stochastic process with even sometimes their baseline activity is very close to a random process (CITE TEODIANO BASTOS).
 
@@
 
+The EEG markers (Lotte again) that can be used to transmit volitionally information are limited, and each one of them has a particular set or combination of methods to decode them.
+adaptative and coadaptation is a must.
+
 In this work a new method to characterize EEG signals is presented, expanded and detailed.
-Its validity is shown by processing offline data from ALS patients.
-This is the continuation of the work previously presented \citep{Ramele2016}, which was applied mainly to rythmic patterns.
+Its validity is verified by processing off-line data for ALS patients.
+This is the continuation of the work previously presented \citep{Ramele2016}, which was previously applied exclusively to rhythmic patterns.
 
@@
 
+This paper reports a straightforward method to, first characterize EEG signals based on the identification of their morphological shapes, and how this characterization can be used to implement a BCI scheme to identify Event Related Potentials, particularly the well-known P300, on an off-line and public dataset.
+
 \section{Materials and Methods}
@@
 Besides new novel applications of BCI, the core value of the entire discipline is to provide communication assistance to patients.
-It is understood that people suffering from ALS, even though the disease do not harm the CNS, their brain signals and suffer slow and deterioriating changes that lead to more complex patterns to be detected [cite]
+It is understood that people suffering from ALS, even though the disease do not harm the CNS, their brain signals and suffer slow and deteriorating changes that lead to more complex patterns to be detected [cite]
 
-Although they are used (PAPER DE P300 donde dice que se usa en los laboratorios) they still lack the bla bla bla CITE LOTTE BOOK \citep{Huggins2016}
+Although this quickly developing discipline has gained mainstream public awareness, held a particular  Olympics \citep{Cybatholn} and even been broadcasted in the last World Cup,  they are used \citep{Huggins2016}) they still lack the necessary robustness, and its performance still lacks well behind any other method of human computer interaction, including any kind of remaining movement \citep{Clerc}.
 
@@
 
-BCI Simulation vs k-fold cross validation
+generative
+
+BCI Simulation vs k-fold cross validation.
+It is quite known and understood that because the same experience will not generate the same markers in the EEG, BCI Simulation is sometimes better suited to understand how the device will operate on a real online application (the graeat variability of BCI).
 
@@
 \centering
-\includegraphics[width=18cm]{subjectaverageds4.eps}
+\includegraphics[width=18cm]{subjectaveraged.eps}
 \caption{This is a figure, Schemes follow the same formatting.
@@
 We also achieved single trial classification for just one subject.
+
+Databases of descriptors.  Flexibility of the descriptor in terms of configuration and size (reference variants of SIFT on lower dimensiones).
+
+Visual inspections enhancement tools.
+
+Understanding and expanding the use of this tool to automatically classify patterns in EEG, that are specifically identified by their shapes.
+This tool may provide the ability to help the physician to locate this patterns particularly in long-periods of time or long recordings.
+\citep{Hartman2005}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{Juan Miguel Santos organized the project and conceived the idea. All the authors equally worked in the completion of this project.}
+\authorcontributions{Juan Miguel Santos lead the project and conceived the idea. All the authors equally worked in the completion of this project.}
 

### 2016-12-19 article.tex
Better version

@@
 % Authors, for the paper (add full first names)
-\Author{Rodrigo Ramele $^{1,\dagger,\ddagger}$, Ana Julia Villar $^{1,\ddagger}$ and Juan Miguel Santos  $^{1,\ddagger}$*}
+\Author{Rodrigo Ramele $^{1,\dagger}$, Ana Julia Villar $^{1}$ and Juan Miguel Santos  $^{1}$*}
 % Authors, for metadata in PDF
@@
 \firstnote{Current address: C1437FBH Lavarden 315, Ciudad Autónoma de Buenos Aires, Argentina}
-\secondnote{These authors contributed equally to this work.}
 
@@
 
-Although recently impressive advances in neuroimagining (particularly radio-nuclear and radiological scanning methods) \citep{Schomer2010} techniques has pushed the traditional EEG, Electroencephalography technology a little bit out of the way in clinical context, the advent and advances in digitalized counterparts has pushed for a revamping of the entire discipline.
+Although recently impressive advances in neuroimagining techniques (particularly radio-nuclear and radiological scanning methods) \citep{Schomer2010} has diminished the prospects of the traditional EEG, Electroencephalography, the advent and development in digitalized devices has pushed for a revamping of the entire discipline.
 
-Their versatility, ease of use, temporal resolution, ease of development and fabrication, and the explosion as consumer devices, is pushing EEG devices to become the de-facto non invasive method to access and harness brain information.
+Their versatility, ease of use, temporal resolution, ease of development and fabrication, and its proliferation as consumer devices, is pushing EEG devices to become the de-facto non invasive method to access and harness brain information.
 
-Key contribution to this development has been the field of Brain Computer Interfaces, which is the development of a new channel of communication particularly aimed to person with some degree of neuro degenerative disease.
+Key contribution to this development has been the field of Brain Computer Interfaces (BCI) \citep{Wolpaw Seminar paper} which is the pursuit of the development of a new channel of communication particularly aimed to persons affected by neuro-degenerative  diseases.
 
-One very remarkable aspect of this communication channel is the ability to volitionally transmit information from the CNS, Central Nervous System, to a computer device and from there use that information to drive wheelchair or as input in a speller application \citep{Guger2009a}, or a rehabilitation procedure \citep{Jure2016}.
-The golden goal of BCI is to implement a new complete and alternative pathway to restore motion \citep{WolpawJonathanR2012}.
+One very noteworthy aspect of this communication channel is the ability to volitionally transmit information from the CNS, Central Nervous System, to a computer device and from there use that information to control a wheelchair \citep{Carlson2013}, as input to a speller application \citep{Guger2009a}, or as aiding tool in a rehabilitation procedure \citep{Jure2016}.
+The holly grail of BCI is to implement a new complete and alternative pathway to restore lost locomotion \citep{WolpawJonathanR2012}.
 
-EEG signals are remarkably complex and has been characterized as a non-stationary stochastic process with even sometimes their baseline activity is very close to a random process (CITE TEODIANO BASTOS).
-
-Additionally, they have a high variability between different subjects and even between different moments for the same subject.
-This inherent complexity is a real challenge when it is required to extract information from raw EEG signals.
+EEG signals are remarkably complex and have been characterized as a multichannel non-stationary stochastic process and even considered as sufficiently random as to be considered as a source of pseudo random number generator \citep{Chen2014}.
+Additionally, they have a high variability between different subjects and even between different moments for the same subject, requiring an adaptive and co-adaptive calibration and learning procedure \citep{Clerc}.
+This imposes an outstanding challenge that is necessary to overcome in order to extract information from raw EEG signals.
 
@@
 
-The EEG markers (Lotte again) that can be used to transmit volitionally information are limited, and each one of them has a particular set or combination of methods to decode them.
-adaptative and coadaptation is a must.
+EEG markers \citep{Clerc} that can be used to transmit volitionally information are limited, and each one of them has a particular set or combination of methods to decode them.
 
@@
 Its validity is verified by processing off-line data for ALS patients.
-This is the continuation of the work previously presented \citep{Ramele2016}, which was previously applied exclusively to rhythmic patterns.
+This is the continuation of the work previously presented \citep{Ramele2016}, which was applied to rhythmic patterns.
 
-The same methodological algorithm can be used to process rhythmic EEG markers or paradigms and at the same time to analyze spatiotemporal phenomena like ERP, particularly P300 \citep{Ramele2016}.
+The same algorithm can be used to process rhythmic EEG markers or paradigms and at the same time to analyze spatiotemporal phenomena like ERP, particularly P300 \citep{Ramele2016}.
 
-This method is based on a morphological analysis of the shape of the EEG signal.
+This method is based on a morphological analysis of the shape of the EEG signal and was inspired by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting the raw signals.
 
-This paper reports on a study done to address
-the hypothesis that communication task complexity
-(copy-spelling vs. novel text generation of a picture
-description) is correlated with BCI performance accuracy
-and P300 characteristics.
+%This paper reports on a study done to address
+%the hypothesis that communication task complexity
+%(copy-spelling vs. novel text generation of a picture
+%description) is correlated with BCI performance accuracy
+%and P300 characteristics.
 
@@
 
-Besides new novel applications of BCI, the core value of the entire discipline is to provide communication assistance to patients.
-It is understood that people suffering from ALS, even though the disease do not harm the CNS, their brain signals and suffer slow and deteriorating changes that lead to more complex patterns to be detected [cite]
+Besides new novel applications of BCI, the core value of the entire discipline is to provide communication assistance to people affected by neuro-degenerative diseases.
+It is understood that people suffering from ALS, even though the disease do not harm the CNS, their brain signals do suffer slow and deteriorating changes that lead to different EEG  patterns\citep{Riener2014}.
 
-Although this quickly developing discipline has gained mainstream public awareness, held a particular  Olympics \citep{Cybatholn} and even been broadcasted in the last World Cup,  they are used \citep{Huggins2016}) they still lack the necessary robustness, and its performance still lacks well behind any other method of human computer interaction, including any kind of remaining movement \citep{Clerc}.
+Although this quickly developing discipline has gained mainstream public awareness, held a peculiar Olympics \citep{Riener2014} and even been broadcasted during the last World Cup,  they are used \citep{Huggins2016}) they still lack the necessary robustness, and its performance still lacks well behind any other method of human computer interaction, including any kind of remaining movement \citep{Clerc}.
 
@@
 
-The methods for averaging.
+The methods for averaging.  \citep{Liang2008}
 
@@
 
-In Fig.
-\ref{fig:subjectaveraged} the grand average (point-to-point) of for all the subjects is shown.
-The P300 characteristic curve can be seen particularly in subjects 6, and 7 and in a lesser extend in the remaining subjects.
-In order to obtain a valid binary classification on averaged signals, particular care was observed to avoid unbalanced samples, because that may introduce a bias in the classification procedure (signal variance is proportional to the root of the number of samples and the procedure would be discriminating signals with different variances).it is quite likely that the data will be unbalanced (that is, an unequal number of trials in each condition)
+In Figure \ref{fig:subjectaveraged} the grand average (point-to-point) for all the subjects is shown.
+The P300 characteristic curve can be seen particularly in subjects 2, and 6 and in a lesser extend in the remaining subjects.
+In order to obtain a valid binary classification on averaged signals, particular care was observed to avoid unbalanced samples, because that may introduce a bias in the classification procedure (signal variance is proportional to the square root of the number of samples and the procedure would be discriminating signals with different variances).Particularly in the case of the P300 response, the oddball paradigm requires that one of the stimulus need to be infrequent so that will unavoidably force that the data will be unbalanced (that is, an unequal number of trials in each condition).
 
@@
 \centering
-\includegraphics[width=18cm]{subjectaveraged.eps}
+\includegraphics[width=16cm]{subjectaveraged.eps}
 \caption{This is a figure, Schemes follow the same formatting.
@@
 
-Results are shown in the following chart where BCI accuracy
-was calculated as the percentage of successful predictions
-that matched the proposed selection in a copy-spelling approach (CITE TAYLOR AND  FRANCOS PAPER).
+Results are shown in Table \ref{tab:results} where BCI accuracy
+for the Cz channel is calculated and compared with the values obtained by the published of the dataset.
+Additionally the best performing channel is informed and its accuracy value.
+It is of particular interest that using this method, the best performing channel was not always Cz, and instead occipital channels PO8 and PO7 showed very good performance indeed \citep{Huggins2016,Jure2016}.
+
+ (CITE TAYLOR AND  FRANCOS PAPER).
 
@@
 %% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
-\begin{tabular}{cccc}
+\begin{tabular}{ccccc}
 \toprule
-\textbf{Cz}	& \textbf{Cz}	& \textbf{Best Channel}	& \textbf{Performance}\\
+\textbf{Participants}	& \textbf{Original ACC at Cz}	& \textbf{ACC at Cz}	& \textbf{Best Performing Channel}	& \textbf{Performance}\\
 \midrule
-0.845000     & 0.794286 & 2 & 0.794286 $\pm$ 0.011952 \\
-0.863000     & 0.811429 & 7 & 0.934286 $\pm$ 0.011200 \\
-0.872000     & 0.780000 & 2 & 0.780000 $\pm$ 0.031843 \\
-0.859000     & 0.675714 & 8 & 0.904286 $\pm$ 0.013232 \\
-0.862000     & 0.801429 & 7 & 0.932857 $\pm$ 0.008400 \\
-0.886000     & 0.964286 & 2 & 0.964286 $\pm$ 0.010435 \\
-0.886000     & 0.784286 & 7 & 0.930000 $\pm$ 0.009744 \\
-0.923000     & 0.908571 & 7 & 0.997143 $\pm$ 0.003733 \\
+1     &   0.84 &   0.79 & Cz  &   0.79 $\pm$ 0.01 \\
+2     &   0.86 &   0.81 & PO7 &   0.93 $\pm$ 0.01 \\
+3     &   0.87 &   0.78 & Cz  &   0.78 $\pm$ 0.03 \\
+4     &   0.86 &   0.68 & PO8 &   0.90 $\pm$ 0.01 \\
+5     &   0.86 &   0.80 & PO7 &   0.93 $\pm$ 0.01 \\
+6     &   0.89 &   0.96 & Cz  &   0.96 $\pm$ 0.01 \\
+7     &   0.89 &   0.78 & PO7 &   0.93 $\pm$ 0.01 \\
+8     &   0.92 &   0.91 & PO7 &   1.00 $\pm$ 0.00 \\
 \bottomrule
 \end{tabular}
+\label{tab:results}
 \end{table}
 
-Contrary to the literature (\citep{Huggins2016,Jure2016}) the best performing channels were not Cz and Pz but instead the occipital channels PO1 and P02 (5 against 3).
-
-The ITR, or BTR, in the case of reactive BCIs (CITE WOLPAW) strongly depends on the amount of averaging required to transmit a valid and robust selection.
-If the amount of repetitions is reduced for each subject, performing a BCI Simulation can be shown on Fig.
-\ref{fig:singletrial}
+The ITR, or BTR, in the case of reactive BCIs \citep{WolpawJonathanR2012} strongly depends on the amount of signal averaging required to transmit a valid and robust selection.
+There is a trade-off that needs to be maintained between the amount of repetitions for each trial  \ref{fig:singletrial}
 
@@
 \centering
-\includegraphics[width=18cm]{singletriality.eps}
+\includegraphics[width=16cm]{singletriality.eps}
 \caption{This is a figure, Schemes follow the same formatting.
@@
 
-\noindent MDPI: Multidisciplinary Digital Publishing Institute\\
-DOAJ: Directory of open access journals\\
-TLA: Three letter acronym\\
+\noindent EEG: Multidisciplinary Digital Publishing Institute\\
+BCI: Directory of open access journals\\
+SNR: Three letter acronym\\
 LD: linear dichroism}

### 2016-12-20 coverletter.tex
First draft version

Dear Editor,

This submitted article encompass the work that we have been doing in a context of a PhD thesis and in the context of the setup of a Brain Computer Interfaces Laboratory at the Computer Engineering Department of the ITBA University in Buenos Aires Argentina.

EEG Processing and BCI are research areas with amazing prospects for the general population but also to very specific population groups that require all the help that we can provide, as engineers and researchers.

We believe that this proposed method based on the histograms of gradients of signal plots, and most importantly, its successful application to identify ERP markers within EEG signals from ALS patients is relevant for this special issue and that it will contribute with a novel tool to analyse EEG signals.

Yours truly and deeply greatfull for your help
Rodrigo Ramele
PhD Candidate
ITBA University 
Argentina

### 2016-12-20 article_vJ.tex
Add files via upload
van los cambios en la sección 1.2.3

%  LaTeX support: latex@mdpi.com 
%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

% You need to save the "mdpi.cls" and "mdpi.bst" files into the same folder as this template file.

%=================================================================
\documentclass[entropy,article,submit,moreauthors,pdftex,10pt,a4paper]{mdpi}

%
%--------------------
% Class Options:
%--------------------
% journal
%----------
% Choose between the following MDPI journals:
% actuators, admsci, aerospace, agriculture, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, atmosphere, atoms, axioms, batteries, behavsci, beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci, buildings, carbon, cancers, catalysts, cells, challenges, chemosensors, children, chromatography, climate, coatings, computation, computers, condensedmatter, cosmetics, cryptography, crystals, data, dentistry, designs, diagnostics, diseases, diversity, econometrics, economies, education, electronics, energies, entropy, environments, epigenomes, fermentation, fibers, fishes, fluids, foods, forests, futureinternet, galaxies, games, gels, genealogy, genes, geosciences, geriatrics, healthcare, horticulturae, humanities, hydrology, informatics, information, infrastructures, inorganics, insects, instruments, ijerph, ijfs, ijms, ijgi, ijtpp, inventions, jcdd, jcm, jdb, jfb, jfmk, jimaging, jof, jintelligence, jlpea, jmse, jpm, jrfm, jsan, land, languages, laws, life, literature, lubricants, machines, magnetochemistry, marinedrugs, materials, mathematics, mca, mti, medsci, medicines, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, molbank, molecules, mps, nanomaterials, ncrna, neonatalscreening, nutrients, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, plants, polymers, processes, proteomes, publications, recycling, religions, remotesensing, resources, risks, robotics, safety, sensors, separations, sexes, sinusitis, socsci, societies, soils, sports, standards, sustainability, symmetry, systems, technologies, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vetsci, viruses, water
%---------
% article
%---------
% The default type of manuscript is article, but can be replaced by: 
% addendum, article, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, correction, conferenceproceedings, conferencereport, expressionofconcern, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimage, letter, newbookreceived, opinion, obituary, projectreport, reply, retraction, review, preprints, shortnote, supfile, technicalnote
% supfile = supplementary materials
%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g. the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figure are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother 
\articlenumber{x}
\doinum{10.3390/------}
\pubvolume{xx}
\pubyear{2016}
\copyrightyear{2016}
\externaleditor{Academic Editor: name}
\history{Received: date; Accepted: date; Published: date}

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Remark, Definition
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{EEG characterization and classification based on histograms of visual gradients}

% Authors, for the paper (add full first names)
\Author{Rodrigo Ramele $^{1,\dagger}$, Ana Julia Villar $^{1}$ and Juan Miguel Santos  $^{1}$*}
% Authors, for metadata in PDF
\AuthorNames{Rodrigo Ramele, Ana Julia Villar and Juan Miguel Santos}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address[1]{%
$^{1}$ \quad Computer Engineering Department, Instituto Tecnológico de Buenos Aires (ITBA); info@itba.edu.ar}

% Contact information of the corresponding author
\corres{Correspondence: rramele@itba.edu.ar; Tel.: +54-9-11-4193-9382}

% Current address and/or shared authorship
\firstnote{Current address: C1437FBH Lavarden 315, Ciudad Autónoma de Buenos Aires, Argentina}

% Simple summary
%\simplesumm{}

% Abstract (Do not use inserted blank lines, i.e. \\) 
\abstract{The analysis of Electroencephalograpy (EEG) signals for Amyotrophic Lateral Sclerosis (ALS) patients is of ulterior importance to elucidate different patterns or markers that could be potentially used to aid in diagnosis or treatment and to implement Brain Computer Interface (BCI) systems which may provide alternative pathways to transmit volitional information to enhance patient's quality of life.  Of particular interests are those which are based on the recognition of Event-Related Potentials (ERP) which are shown to be more robust to the unavoidable physical impairment which comes with this motor neuron disease.  This work mimics what electroencephalographers has been doing clinically, visually inspecting and categorizing phenomena within the EEG, and provides a framework to analyse and characterize EEG signals, with a focus on the P300, an ERP elicited by the oddball paradigm of rare events.  The validity of the method is shown by off-line processing a public dataset of ALS patients.}

% Keywords
\keyword{electroencephalography (EEG); BCI; P300; ALS; classification; HOG; SIFT}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}
%\AMS{}

% If this is an expanded version of a conference paper, please cite it here: enter the full citation of your conference paper, and add $^\S$ in the end of the title of this article.
%\conference{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:

%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% For Conference Proceedings Papers: add the conference title here
%\conferencetitle{}

%\setcounter{secnumdepth}{4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Only for the journal Gels: Please place the Experimental Section after the Conclusions

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{section}{-1} %% Remove this when starting to work on the template.

\section{Introduction}

%\begin{enumerate}[leftmargin=*,labelsep=3mm]
%\item	Why
%\item	Purpose and significance
%\item	Current state of the research field and key publications cited
%\item   Controversial hypothesis
%\item   Main aim and principal conclusions.
%\end{enumerate}
%
%\begin{enumerate}[leftmargin=*,labelsep=3mm]
%\item El paper tiene un foco general en EEG y rapidamente se centra en BCI como una prueba del método.
%\item La relevancia de estudiar métodos de análisis de EEG para pacientes con ALS.
%\item Ofrece una Receta de cómo implementar un esquema simple de detección de P300 y ofrece una comparativa contra el paper de las italianas.
%\item La transversalidad del método de poder detectar eventos rítmicos asi como transientes como p300.
%\item La posibilidad de utilizarlo para reconocer, fuertemente un patron particular y con la flexibilidad de detectar variantes basado en las poblaciones de descriptores (esta es mi hipótesis de porque funciona, porque yo vi que el método se hace pelota con pequeñas variantes de la señal).
%\item La importancia del averaging de EEG y sus desbalanceos, puntualizando que en el paper original
%\item Que la caída en la clasificación no cae significativamente aún bajando la cantidad de repeticiones requeridas para el averaging.
%\item ¿ Por qué da mejor en PO8 que en Cz y Pz que son los canales P300 ?  (como future work).
%\end{enumerate}

Although impressive advances in neuroimagining techniques (particularly radio-nuclear and radiological scanning methods) \citep{Schomer2010} has diminished the prospects of the traditional Electroencephalography (EEG), the advent and development of digitalized devices has pressed for a revamping of this hundredth years old technology.  Their versatility, ease of use, temporal resolution, ease of development and fabrication, and its proliferation as consumer devices, is pushing EEG to become the de-facto non invasive method to access and harness brain information.

Key contribution to this development has been the field of Brain Computer Interfaces (BCI) \citep{WolpawJonathanR2012} which is the pursuit of the development of a new channel of communication particularly aimed to persons affected by neuro-degenerative diseases.

One very noteworthy aspect of this communication channel is the ability to volitionally transmit information from the Central Nervous System (CNS), to a computer device and from there use that information to control a wheelchair \citep{Carlson2013}, as input to a speller application \citep{Guger2009a}, or as aiding tool in a rehabilitation procedure \citep{Jure2016}.  The holly grail of BCI is to implement a new complete and alternative pathway to restore lost locomotion \citep{WolpawJonathanR2012}.

EEG signals are remarkably complex and have been characterized as a multichannel non-stationary stochastic process and even considered random enough as to be used as a source of pseudo random number generator \citep{Chen2014}. Additionally, they have a high variability between different subjects and even between different moments for the same subject, requiring an adaptive and co-adaptive calibration and learning procedures \citep{Clerc}.  Hence, this imposes an outstanding challenge that is necessary to overcome in order to extract information from raw EEG signals.

EEG markers \citep{Clerc} that can be used to transmit volitionally information are limited, and each one of them has a particular set or combination of appropriate methods to decode them. Inevitably, it is often necessary to implement many distinct and specialized algorithmic methods, to filter the signal, enhance its SNR, and try to determine some meaning out of it.

In this work a new method to characterize EEG signals is presented, expanded and detailed.  Its validity is verified by processing off-line data for ALS patients.  This is the continuation of the work previously presented in \citep{Ramele2016}, where it was applied to rhythmic patterns, and that it can be extended to describe transient events like those produced by the P300\citep{Knuth2006}.

The method is based on the morphological analysis of the shape of the EEG signal\citep{Alvarado-Gonzalez2016,Yamaguchi2009} and was inspired by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting  raw signals.

%This paper reports on a study done to address
%the hypothesis that communication task complexity
%(copy-spelling vs. novel text generation of a picture
%description) is correlated with BCI performance accuracy
%and P300 characteristics.

In brief, this paper reports a straightforward method to, first characterize EEG signals based on the identification of their morphological shapes, and how this characterization can be used to implement a BCI scheme to identify Event Related Potentials, particularly the well-known P300, on an off-line and public dataset.

\section{Materials and Methods}

Besides new novel applications of BCI, the goal of the entire discipline is to provide communication assistance to people affected by neuro-degenerative diseases.  It is understood that for people suffering from ALS, even though this pathology do not harm the CNS, their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns \citep{Nijboer2009,Riener2014}.

BCI has gained mainstream public awareness and recently even held a peculiar Olympics \citep{Riener2014} and been broadcasted during the inauguration of the last World Cup.  New developments have overcome the out-of-the-lab high-bar and they are starting to be used in real world environments\citep{Huggins2016}.  However, they still lack the necessary robustness, and its performance is well behind any other method of human computer interaction, including any kind of detection of remaining muscular movement \citep{Clerc}.

To verify the validity of the proposed framework and method, the public dataset 008-2014 published on the BNCI-Horizon website \citep{Riccio2013} was used to perform a binary classification task on the provided signals.  The algorithm was implemented using the VLFeat \citep{Vedaldi2010} Computer Vision libraries on MATLAB 2014a (Mathworks Inc., Natick, MA, USA).

\subsection{Experimental protocol}

The protocol is explained in \citep{Riccio2013} but can be summarized as follows:  8 subjects with confirmed diagnoses but on different stages of ALS disease, were recruited and accepted to perform the experiments. The P300 task objective was meant to spell 7 words of 5 letter each.  The first 3 runs were used for training and the remaining 4 for testing with visual feedback.  A trial, for this experiment, was defined as every attempt to select a letter from the speller, and it was composed of 10 repetitions of flashes of 6 rows and 6 columns of the traditional 6x6 P300 matrix , yielding 120 repetitions.  Flashing of row/columns was performed for 0.125 s, following by a resting period of the same size.  After 120 repetions and inter-trial pause was included before resuming with the following letter.

The gathered dataset was sampled at 256 Hz and consisted of EEG matrix for electrode channels Fz,Cz,Pz,Oz,P3,P4,PO7 and PO8, identified according to the 10-20 International System,  for each one of the 8 subjects.  This dataset was published in the popular BNCI-Horizon 2020 website \citep{Brunner2014}.

%\subsection{Morphological Signal Analysis}
%
%Several approaches and methods have been applied to decode them.  Of particular interests are morphological signal analysis 
%CITE (Paper del chabon del ITBA que detalla bien los diferentes metodos) \citep{Alvarado-Gonzalez2016,Yamaguchi2009}.
%
%Morphological or 
%
%template matching
%
%spike
%sharp wave
%spike wave
%polyspike 
%slow wave complex
%
%shape domain
%
%\subsection{P300}
%
%Uhhhh I can talk a lot about P300 \citep{Knuth2006}.
%
%P300 single trial is a golden grail of Brain computer interfaces.  It is studied and analyzed in this way, here in this other way, in this other way is here.

\subsection{Algorithm}

\subsubsection{Preprocessing}

The mandatory first step is to help to enhance the SNR of the P300 pattern above the level of basal EEG. The processing pipeline starts with applying a notch filter to the raw signal, a decimation with a FIR filter of order 30 from the original 256 Hz to 32 Hz and finally applying a 4th degree 10 Hz lowpass Butterworth filter.

\subsubsection{Segmentation, Artefact Removal and Signal Averaging}

The EEG matrix is processed on a channel by channel basis.  For each word, each letter, a segment of EEG signal is extracted that encompass 10 repetitions of 12 flashings, 6 rows and 6 columns.  For every 12 stimuli flashing, i.e. one complete sequence of repetitions of each row/column, a basic artefact removal procedure of removing the entire segment if any signal goes above 70 $\mu$V is applied.

From every segment of 12 flashings, two epoch corresponding to a hit and two matching a no-hit are constructed as segments of 1s length (256 sample points).

These segments, 20 for each class, are later point-to-point averaged for the whole 120 repetitions group.  This finally lead to 2 averaged signals for each of the 5 letters, for the 7 words, comprising a set of 70 averaged signals\citep{Liang2008}.

\subsubsection{Features}

The central part of this method is the feature generation: the histogram of visual gradients.  To do so, the first step is the transformation of the signal into a temporary binary image.

The signal is first scaled and centered at zero by

\begin{equation}
\tilde{x}(t,c) = \left \lfloor{ \gamma \cdot ( x(t,c) - \bar{x}(t,c)  )}\right \rfloor
\end{equation}

\noindent where $\gamma$ is the image scale, $t$ is time and $ x(t,c) $ is the EEG matrix defined for each $t$ and for a particular channel $c$.

Then the image is constructed by

\begin{equation}
I(z_1,z_2) = \left\{ \begin{array}{rl}
255 & z_1 = \gamma \cdot t; z_2 = \tilde{x}(t,c) + z(c) \\
0   & \mbox{otherwise}
\end{array}\right.
\label{eq:images}
\end{equation}

\noindent and used to derive local representations that will capture the visual shape of the signal.

Finally, the histogram of gradients, which is based on a technique from Computer Vision, the SIFT \citep{Lowe2004} descriptor, is calculated on local image patches centered on sample points \citep{Vedaldi2010}.
To do this, the patch is divided in 16 blocks, arranged in a 4x4 grid and centered on $T = (x,y)$, a location within the image.  On each block, the bin consists of 45 degrees obtaining 8 aligning angles.

The values of the image gradients aligned with each angle bin are accumulated through

\begin{equation}
 h(p,i,j) = 3 \cdot s \int w_\mathrm{ang}(\angle J(\mathbf{x}) - \theta_p)\, w_{ij}\left(\frac{\mathbf{x} - T}{3 s}\right)\, |J(\mathbf{x})|\, d\mathbf{x}
\label{eq:histogram}
\end{equation}

\noindent where $s$ is the size of the local patch, $\angle J(\mathbf{x})$ is the gradient found at each of the 16 blocks of the patch aligned with the angle $\theta_p$.  
This gives a descriptor of 128 dimension as shown on Figure \ref{fig:sampledescriptor}.

\begin{figure}[H]
\centering
\includegraphics[width=16cm]{sampledescriptor.eps}
\caption{The local patch is located around a sample point plotted on the temporary signal. All the sample points are interpolated using the Bresenham algorithm. The 4x4 grid can be seen on the descriptor and also arrows showing the dominant gradient on each block.}
\label{fig:sampledescriptor}
\end{figure}

\subsubsection{Classification}

The classification is carried out by using the discriminative semi-supervised classification method, the Naive Bayes Nearest Neighbout (NBNN) \citep{Boiman2008}. % algorithm which can be described as follows.

\begin{equation}
\hat{C} = argmin_C \sum_{}^{} \left\lVert d_i - NN_C(d_i) \right\rVert ^2
\end{equation}

%BCI Simulation vs k-fold cross validation.  It is quite known and understood that because the same experience will not generate the same markers in the EEG, BCI Simulation is sometimes better suited to understand how the device will operate on a real online application (the graeat variability of BCI).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

In Figure \ref{fig:subjectaveraged} the grand average (point-to-point) for all the subjects is shown.  The P300 characteristic curve can be seen particularly in subjects 2, and 6 and in a lesser extend in the remaining subjects. In order to obtain a valid binary classification on averaged signals, particular care was observed to avoid unbalanced samples, because that may introduce a bias in the classification procedure (signal variance is proportional to the square root of the number of samples and the procedure would be discriminating signals with different variances).Particularly in the case of the P300 response, the oddball paradigm requires that one of the stimulus need to be infrequent so that will unavoidably force that the data will be unbalanced (that is, an unequal number of trials in each condition).

\begin{figure}[H]
\centering
\includegraphics[width=16cm]{subjectaveraged.eps}
\caption{Point-To-Point grand averages of epochs obtained for hits (solid line) and no hits (dashed line) for each one of the 8 subjects for channel Cz. The P300 characteristic curve can be well identified on subjects 2 and 6.}
\label{fig:subjectaveraged}
\end{figure}

Results are shown in Table \ref{tab:results} where BCI accuracy
for the Cz channel is calculated and compared with the values obtained by the published of the dataset.  Additionally the best performing channel is informed and its accuracy value.  It is of particular interest that using this method, the best performing channel was not always Cz, and instead occipital channels PO8 and PO7 showed very good performance indeed \citep{Huggins2016,Jure2016}.

% (CITE TAYLOR AND  FRANCOS PAPER).

\begin{table}[H]
\caption{Results obtained for Accuracy levels by 3-fold cross validation. The values reported by the dataset publishers for Cz are reproduced here for comparison. Additionally, channel Cz BCI accuracies can be seen as well as the performance levels obtained for the BPC, the best performing channel for each subject.}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{ccccc}
\toprule
\textbf{Participants}	& \textbf{Original Cz}	& \textbf{ACC at Cz}	& \textbf{BPC}	& \textbf{Performance}\\
\midrule
1     &   0.84 &   0.79 & Cz  &   0.79 $\pm$ 0.01 \\
2     &   0.86 &   0.81 & PO7 &   0.93 $\pm$ 0.01 \\
3     &   0.87 &   0.78 & Cz  &   0.78 $\pm$ 0.03 \\
4     &   0.86 &   0.68 & PO8 &   0.90 $\pm$ 0.01 \\
5     &   0.86 &   0.80 & PO7 &   0.93 $\pm$ 0.01 \\
6     &   0.89 &   0.96 & Cz  &   0.96 $\pm$ 0.01 \\
7     &   0.89 &   0.78 & PO7 &   0.93 $\pm$ 0.01 \\
8     &   0.92 &   0.91 & PO7 &   1.00 $\pm$ 0.00 \\
\bottomrule
\end{tabular}
\label{tab:results}
\end{table}

The ITR, or BTR, in the case of reactive BCIs \citep{WolpawJonathanR2012} strongly depends on the amount of signal averaging required to transmit a valid and robust selection.  There is a trade-off that needs to be balanced between the required number of repetitions for each trial to guarantee robust transmission and the the achieved speed of transmission affected by the repetions.  As detailed in the table \ref{tab:singletrialreduction} this method is only reduced in a 20 percent in accuracy levels even by using only one sequence of repetitions.

\begin{table}[H]
\caption{Results obtained for Accuracy levels by performing a BCI simulation for the testing dataset. The BPC, best performing channel is shown and also performance levels for the 120, 60 and 12 repetitions. Finally the PR, performance reduction is described.  Almost all the subjects performed around 70\% even with by averaging 2 signals for each class.}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{cccccc}
\toprule
\textbf{Participants}	& \textbf{BPC}	& \textbf{ACC@120}	& \textbf{ACC@60}	&  \textbf{ACC@12} & \textbf{PR}\\
\midrule
1 & Cz &   0.80 &   0.82 &   0.72 & 9\% \\
2 & PO7 &   0.93 &   0.78 &   0.75 & 18\%\\
3 & Cz &   0.80 &   0.82 &   0.72 & 9\%\\
4 & PO8 &   0.93 &   0.78 &   0.65 & 29\%\\
5 & PO7 &   0.93 &   0.78 &   0.75 & 18\%\\
6 & Cz &   0.80 &   0.82 &   0.72 & 9\%\\
7 & PO7 &   0.93 &   0.78 &   0.75 & 18\%\\
8 & PO7 &   0.93 &   0.78 &   0.75 & 18\%\\
\bottomrule
\end{tabular}
\label{tab:singletrialreduction}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Discussion}
%
%This method, different from other methods which is based on the nonlinearity of the gradient of histograms which can be used to detect 
%
%is also based on how the image look like.
%
%SNR of p300 and how to detect it
%
%Check if you can use this to detect any kind of transient signal.
%
%Compare if it is possible with the descriptors from one subject, discriminate the others.
%
%Channal identification based on the metric distance between the bags

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

A method to characterize and classify EEG signals where the main characteristic can be both, rhythmic in nature as in motor imagery, and also transient in time space, like the P300 has been presented.

Although single trial classification was not perfectly achieved, it has been shown that this method could be applied with a P300 Speller Matrix application with increased ITR.

%Databases of descriptors.  Flexibility of the descriptor in terms of configuration and size (reference variants of SIFT on lower dimensiones).

The adaptive behaviour of the algorithm make is well suited when the shape of the pattern elicited by the P300 response, does not conform to the predicted structure.  This is because, the descriptors are directly based on how the signals actually looked like for the training and calibration step, and they do not require any prior knowledge about the signal.  This is of particular relevance for studies on outliers populations as may be the case for people who suffered some neuro-degenerative disease like Lou Gehrig's disease.

The expanding of the understanding of this tool in order to automatically classify patterns in EEG, that are specifically identified by their shapes, is a prospect future work to be considered.  It may also provide  assistance to physician or electroencephalographers to help them locate these EEG patterns particularly in long recording periods \citep{Hartman2005}, frequent in sleep research.

Moreover, this method can be used as an alternate \textit{BCI predictor} \citep{Clerc} or as a tool for artefact removal (which is performed on many occasions by visually inspecting the signal).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Subsection}
%
%\subsubsection{Subsubsection}
%
%Bulleted lists look like this:
%\begin{itemize}[leftmargin=*,labelsep=4mm]
%\item	First bullet
%\item	Second bullet
%\item	Third bullet
%\end{itemize}
%
%Numbered lists can be added as follows:
%\begin{enumerate}[leftmargin=*,labelsep=3mm]
%\item	First item
%\item	Second item
%\item	Third item
%\end{enumerate}
%
%The text continues here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\supplementary{The following are available online at www.mdpi.com/link, Figure S1: title, Table S1: title, Video S1: title.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{This project was supported by the ITBACyT-15 funding program issued by ITBA University.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{This projects is part of a the first author's PhD Thesis which is directed by Juan Miguel Santos and codirected by Ana Julia Villar.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictofinterests{The authors declare no conflict of interest.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\abbreviations{The following abbreviations are used in this manuscript:\\

\noindent EEG: electroencephalography\\
BCI: Brain Computer Interfaces\\
SNR: Signal to Noise Ratio\\
CNS: Central Nervous System\\
ALS: Amyotrophic Lateral Sclerosis\\
ERP: Event-Related Potential\\
P300: Positive deflection of an Event-Related Potential which occurs 300 ms after onset of stimuli\\
ITR: Information Transfer Rate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
%\appendixsections{multiple} %Leave argument "multiple" if there are multiple sections. Then a counter is printed ("Appendix A?). If there is only one appendix section then change the argument to ?one? and no counter is printed (?Appendix?).
%\appendix
%\section{}
%The appendix is an optional section that can contain details and data supplemental to the main text. For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data. Mathemtaical proofs of results not central to the paper can be added as an appendix.
%
%\section{}
%All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here. 
\bibliographystyle{mdpi}

%=====================================
% References, variant A: internal bibliography
%=====================================
%\renewcommand\bibname{References}
%\begin{thebibliography}{999}
% Reference 1
%\bibitem{ref-journal}
%Lastname, F.; Author, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142-149.
% Reference 2
%\bibitem{ref-book}
%Lastname, F.F.; Author, T. The title of the cited contribution. In {\em The Book Title}; Editor, F., Meditor, A., Eds.; Publishing House: City, Country, 2007; pp. 32-58.
%\end{thebibliography}

%=====================================
% References, variant B: external bibliography
%=====================================
\bibliography{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\sampleavailability{Samples of the compounds ...... are available from the authors.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

### 2016-12-20 article.tex
Added ApeBrainStorm changes.

@@
 % Full title of the paper (Capitalized)
-\Title{EEG characterization and classification based on the histograms of visual gradients of plots}
+\Title{EEG characterization and classification based on histograms of visual gradients}
 
@@
 % Abstract (Do not use inserted blank lines, i.e. \\) 
-\abstract{A single paragraph of about 200 words maximum.
-For research articles, abstracts should give a pertinent overview of the work.
-We strongly encourage authors to use the following style of structured abstracts, but without headings: 1) Background: Place the question addressed in a broad context and highlight the purpose of the study; 2) Methods: Describe briefly the main methods or treatments applied; 3) Results: Summarize the article's main findings; and 4) Conclusion: Indicate the main conclusions or interpretations.
-The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.This is great.
-We had tremendous results.
-Our results are so great that nobody will believe that we actually managed to have them !}
+\abstract{The analysis of Electroencephalograpy (EEG) signals for Amyotrophic Lateral Sclerosis (ALS) patients is of ulterior importance to elucidate different patterns or markers that could be potentially used to aid in diagnosis or treatment.
+Additionally, it will improve the implementation of Brain Computer Interface (BCI) systems which may provide alternative pathways to transmit volitional information to enhance patient's quality of life.
+Of particular interests are those which are based on the recognition of Event-Related Potentials (ERP) which are shown to be more robust to the unavoidable physical impairment which comes with this motor neuron disease.
+This work mimics what electroencephalographers have been doing clinically, visually inspecting and categorizing phenomena within the EEG, and provides a framework to analyse and characterize EEG signals, with a focus on the P300, an ERP elicited by the oddball paradigm of rare events.
+The validity of the method is shown by off-line processing a public dataset of ALS patients.}
 
@@
 
-\begin{enumerate}[leftmargin=*,labelsep=3mm]
-\item	Why
-\item	Purpose and significance
-\item	Current state of the research field and key publications cited
-\item   Controversial hypothesis
-\item   Main aim and principal conclusions.
-\end{enumerate}
+%\begin{enumerate}[leftmargin=*,labelsep=3mm]
+%\item	Why
+%\item	Purpose and significance
+%\item	Current state of the research field and key publications cited
+%\item   Controversial hypothesis
+%\item   Main aim and principal conclusions.
+%\end{enumerate}
+%
+%\begin{enumerate}[leftmargin=*,labelsep=3mm]
+%\item El paper tiene un foco general en EEG y rapidamente se centra en BCI como una prueba del método.
+%\item La relevancia de estudiar métodos de análisis de EEG para pacientes con ALS.
+%\item Ofrece una Receta de cómo implementar un esquema simple de detección de P300 y ofrece una comparativa contra el paper de las italianas.
+%\item La transversalidad del método de poder detectar eventos rítmicos asi como transientes como p300.
+%\item La posibilidad de utilizarlo para reconocer, fuertemente un patron particular y con la flexibilidad de detectar variantes basado en las poblaciones de descriptores (esta es mi hipótesis de porque funciona, porque yo vi que el método se hace pelota con pequeñas variantes de la señal).
+%\item La importancia del averaging de EEG y sus desbalanceos, puntualizando que en el paper original
+%\item Que la caída en la clasificación no cae significativamente aún bajando la cantidad de repeticiones requeridas para el averaging.
+%\item ¿ Por qué da mejor en PO8 que en Cz y Pz que son los canales P300 ?  (como future work).
+%\end{enumerate}
 
-\begin{enumerate}[leftmargin=*,labelsep=3mm]
-\item El paper tiene un foco general en EEG y rapidamente se centra en BCI como una prueba del método.
-\item La relevancia de estudiar métodos de análisis de EEG para pacientes con ALS.
-\item Ofrece una Receta de cómo implementar un esquema simple de detección de P300 y ofrece una comparativa contra el paper de las italianas.
-\item La transversalidad del método de poder detectar eventos rítmicos asi como transientes como p300.
-\item La posibilidad de utilizarlo para reconocer, fuertemente un patron particular y con la flexibilidad de detectar variantes basado en las poblaciones de descriptores (esta es mi hipótesis de porque funciona, porque yo vi que el método se hace pelota con pequeñas variantes de la señal).
-\item La importancia del averaging de EEG y sus desbalanceos, puntualizando que en el paper original
-\item Que la caída en la clasificación no cae significativamente aún bajando la cantidad de repeticiones requeridas para el averaging.
-\item ¿ Por qué da mejor en PO8 que en Cz y Pz que son los canales P300 ?  (como future work).
-\end{enumerate}
+Although impressive advances in neuroimagining techniques (particularly radio-nuclear and radiological scanning methods) \citep{Schomer2010} have diminished the prospects of the traditional Electroencephalography (EEG), the advent and development of digitalized devices has pressed for a revamping of this hundredth years old technology.
+Their versatility, ease of use, temporal resolution, ease of development and fabrication, and its proliferation as consumer devices, are pushing EEG to become the de-facto non invasive method to access and harness brain information.
 
-Although recently impressive advances in neuroimagining techniques (particularly radio-nuclear and radiological scanning methods) \citep{Schomer2010} has diminished the prospects of the traditional EEG, Electroencephalography, the advent and development in digitalized devices has pushed for a revamping of the entire discipline.
+Key contribution to this development has been the field of Brain Computer Interfaces (BCI) \citep{WolpawJonathanR2012} which is the pursuit of the development of a new channel of communication particularly aimed to persons affected by neuro-degenerative diseases.
 
-Their versatility, ease of use, temporal resolution, ease of development and fabrication, and its proliferation as consumer devices, is pushing EEG devices to become the de-facto non invasive method to access and harness brain information.
-
-Key contribution to this development has been the field of Brain Computer Interfaces (BCI) \citep{Wolpaw Seminar paper} which is the pursuit of the development of a new channel of communication particularly aimed to persons affected by neuro-degenerative  diseases.
-
-One very noteworthy aspect of this communication channel is the ability to volitionally transmit information from the CNS, Central Nervous System, to a computer device and from there use that information to control a wheelchair \citep{Carlson2013}, as input to a speller application \citep{Guger2009a}, or as aiding tool in a rehabilitation procedure \citep{Jure2016}.
+One very noteworthy aspect of this novel communication channel is the ability to volitionally transmit information from the Central Nervous System (CNS) to a computer device and from there use that information to control a wheelchair \citep{Carlson2013}, as input to a speller application \citep{Guger2009a}, or as aiding tool in a rehabilitation procedure \citep{Jure2016}.
 The holly grail of BCI is to implement a new complete and alternative pathway to restore lost locomotion \citep{WolpawJonathanR2012}.
 
-EEG signals are remarkably complex and have been characterized as a multichannel non-stationary stochastic process and even considered as sufficiently random as to be considered as a source of pseudo random number generator \citep{Chen2014}.
-Additionally, they have a high variability between different subjects and even between different moments for the same subject, requiring an adaptive and co-adaptive calibration and learning procedure \citep{Clerc}.
-This imposes an outstanding challenge that is necessary to overcome in order to extract information from raw EEG signals.
+EEG signals are remarkably complex and have been characterized as a multichannel non-stationary stochastic process and even considered random enough as to be used as a source of pseudo random number generator \citep{Chen2014}.
+Additionally, they have a high variability between different subjects and even between different moments for the same subject, requiring an adaptive and co-adaptive calibration and learning procedures \citep{Clerc}.
+Hence, this imposes an outstanding challenge that is necessary to overcome in order to extract information from raw EEG signals.
 
-Hence, inevitably, it is often necessary to implement many distinct and specialized algorithmic methods, to filter the signal, enhance its SNR, and try to determine some meaning out of it.
-
-EEG markers \citep{Clerc} that can be used to transmit volitionally information are limited, and each one of them has a particular set or combination of methods to decode them.
+EEG markers \citep{Clerc} that can be used to transmit volitionally information are limited, and each one of them has a particular set or combination of appropriate methods to decode them.
+Inevitably, it is often necessary to implement many distinct and specialized algorithmic methods, to filter the signal, enhance its SNR, and try to determine some meaning out of it.
 
@@
 Its validity is verified by processing off-line data for ALS patients.
-This is the continuation of the work previously presented \citep{Ramele2016}, which was applied to rhythmic patterns.
+This is the continuation of the work previously presented in \citep{Ramele2016}, where it was applied to rhythmic patterns, and that it can be extended to describe transient events like those produced by the P300\citep{Knuth2006}.
 
-The same algorithm can be used to process rhythmic EEG markers or paradigms and at the same time to analyze spatiotemporal phenomena like ERP, particularly P300 \citep{Ramele2016}.
-
-This method is based on a morphological analysis of the shape of the EEG signal and was inspired by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting the raw signals.
+The method is based on the morphological analysis of the shape of the EEG signal\citep{Alvarado-Gonzalez2016,Yamaguchi2009} and was inspired by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting  raw signals.
 
@@
 
-This paper reports a straightforward method to, first characterize EEG signals based on the identification of their morphological shapes, and how this characterization can be used to implement a BCI scheme to identify Event Related Potentials, particularly the well-known P300, on an off-line and public dataset.
+In brief, this paper reports a straightforward method to, first characterize EEG signals based on the identification of their morphological shapes, and how this characterization can be used to implement a BCI scheme to identify Event Related Potentials, particularly the well-known P300, on an off-line and public dataset.
 
@@
 
-Besides new novel applications of BCI, the core value of the entire discipline is to provide communication assistance to people affected by neuro-degenerative diseases.
-It is understood that people suffering from ALS, even though the disease do not harm the CNS, their brain signals do suffer slow and deteriorating changes that lead to different EEG  patterns\citep{Riener2014}.
+Besides new novel applications of BCI, the goal of the entire discipline is to provide communication assistance to people affected by neuro-degenerative diseases.
+It is understood that for people suffering from ALS, even though this pathology do not harm the CNS, their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns \citep{Nijboer2009,Riener2014}.
 
-Although this quickly developing discipline has gained mainstream public awareness, held a peculiar Olympics \citep{Riener2014} and even been broadcasted during the last World Cup,  they are used \citep{Huggins2016}) they still lack the necessary robustness, and its performance still lacks well behind any other method of human computer interaction, including any kind of remaining movement \citep{Clerc}.
+BCI has gained mainstream public awareness and recently even held a peculiar Olympics \citep{Riener2014} and been broadcasted during the inauguration of the last Soccer World Cup.
+New developments have overcome the out-of-the-lab high-bar and they are starting to be used in real world environments\citep{Huggins2016}.
+However, they still lack the necessary robustness, and its performance is well behind any other method of human computer interaction, including any kind of detection of remaining muscular movement \citep{Clerc}.
 
-We use the dataset 008-2014 published on the BNCI-Horizon website.
-This consortium of companies and universities were developed under the umbrella of the EU to define, provide guidance for BCI development \citep{Riccio2013}.
+To verify the validity of the proposed framework and method, the public dataset 008-2014  \citep{Riccio2013} published on the BNCI-Horizon website \citep{Brunner2014} was used to perform a binary classification task on the provided signals.
+The algorithm was implemented using the VLFeat \citep{Vedaldi2010} Computer Vision libraries on MATLAB 2014a (Mathworks Inc., Natick, MA, USA).
 
@@
 
-The protocol is explained \citep{Riccio2013} but can be summarized as follows:  8 subjects affected by different levels of ALS disease, were used to generate all the data.
-The P300 task was meant to spell 7 words of 5 letter each.
+The protocol is explained in \citep{Riccio2013} but can be summarized as follows:  8 subjects with confirmed diagnoses but on different stages of ALS disease, were recruited and accepted to perform the experiments.
+The P300 task objective was meant to spell 7 words of 5 letter each.
 The first 3 runs were used for training and the remaining 4 for testing with visual feedback.
-A trial, for this experiment, was defined as each attempt to select a letter from the speller, and it was composed of 10 repetitions of flashes of 6 rows and 6 columns of the matrix, yielding 120 repetitions.
+A trial, for this experiment, was defined as every attempt to select a letter from the speller, and it was composed of 10 repetitions of flashes of 6 rows and 6 columns of the traditional 6x6 P300 matrix , yielding 120 repetitions.
 Flashing of row/columns was performed for 0.125 s, following by a resting period of the same size.
-After 120 repetions and intertrial pause is performed to start with the following letter.
+After 120 repetions and inter-trial pause was included before resuming with the following letter.
 
-All sessions were conducted on the same day with some minutes break in-between. Typing was done at an average speed of 1 key per second.
+The gathered dataset was sampled at 256 Hz and consisted of EEG matrix for electrode channels Fz,Cz,Pz,Oz,P3,P4,PO7 and PO8, identified according to the 10-20 International System,  for each one of the 8 subjects.
 
-\subsection{Morphological Signal Analysis}
-
-Due to the inner complexity of EEG signals, several approaches and methods have been applied to decode them.  Of particular interests are morphological signal analysis 
-CITE (Paper del chabon del ITBA que detalla bien los diferentes metodos) \citep{Alvarado-Gonzalez2016}.
-
-Morphological or
-
-template matching
-
-spike
-sharp wave
-spike wave
-polyspike 
-slow wave complex
-
-shape domain
-
-\subsubsection{P300}
-
-Uhhhh I can talk a lot about P300 \citep{Knuth2006}.
-
-P300 single trial is a golden grail of Brain computer interfaces.  It is studied and analyzed in this way, here in this other way, in this other way is here.
+%\subsection{Morphological Signal Analysis}
+%
+%Several approaches and methods have been applied to decode them.  Of particular interests are morphological signal analysis 
+%CITE (Paper del chabon del ITBA que detalla bien los diferentes metodos) \citep{Alvarado-Gonzalez2016,Yamaguchi2009}.
+%
+%Morphological or 
+%
+%template matching
+%
+%spike
+%sharp wave
+%spike wave
+%polyspike 
+%slow wave complex
+%
+%shape domain
+%
+%\subsection{P300}
+%
+%Uhhhh I can talk a lot about P300 \citep{Knuth2006}.
+%
+%P300 single trial is a golden grail of Brain computer interfaces.  It is studied and analyzed in this way, here in this other way, in this other way is here.
 
@@
 
-\subsubsection{Signal to Image Transformation}
+The mandatory first step is to enhance the SNR of the P300 pattern above the level of basal EEG.
+The processing pipeline starts with applying a notch filter to the raw signal, a decimation with a FIR filter of order 30 from the original 256 Hz to 32 Hz and finally applying a 4th degree 10 Hz lowpass Butterworth filter.
 
-There are many ways an EEG matrix can be transformed into an image.
+\subsubsection{Segmentation, Artefact Removal and Signal Averaging}
 
-What we do is take the signal and plot it.  Thats all folks.
+The EEG matrix is processed on a channel by channel basis.
+For each word, each letter, a segment of EEG signal is extracted that encompass 10 repetitions of 12 flashings, 6 rows and 6 columns.
+For every 12 stimuli flashing, i.e.
+one complete sequence of repetitions of each row/column, a basic artefact removal procedure of removing the entire segment if any signal goes above 70 $\mu$V is applied.
 
-\subsubsection{Signal Averaging}
+From every segment of 12 flashings, two epoch corresponding to a hit and two matching a no-hit are constructed as segments of 1s length (256 sample points).
 
-The methods for averaging.  \citep{Liang2008}
+These segments, 20 for each class, are later point-to-point averaged for the whole 120 repetitions group.
+This finally lead to 2 averaged signals for each of the 5 letters, for the 7 words, comprising a set of 70 averaged signals\citep{Liang2008}.
+
+\subsubsection{Features}
+
+The central part of this method is the feature generation: the histogram of visual gradients.  To do so, the first step is the transformation of the signal into a temporary binary image.
+
+The signal is first scaled and centered at zero by
+
+\begin{equation}
+\tilde{x}(t,c) = \left \lfloor{ \gamma \cdot ( x(t,c) - \bar{x}(t,c)  )}\right \rfloor
+\end{equation}
+
+\noindent where $\gamma$ is the image scale, $t$ is time and $ x(t,c) $ is the EEG matrix defined for each $t$ and for a particular channel $c$.
+
+Then the image is constructed by
+
+\begin{equation}
+I(z_1,z_2) = \left\{ \begin{array}{rl}
+255 & z_1 = \gamma \cdot t; z_2 = \tilde{x}(t,c) + z(c) \\
+0   & \mbox{otherwise}
+\end{array}\right.
+\label{eq:images}
+\end{equation}
+
+\noindent and used to derive local representations that will capture the visual shape of the signal.
+
+Finally, the histogram of gradients, which is based on a technique from Computer Vision, the SIFT \citep{Lowe2004} descriptor, is calculated on local image patches centered on sample points \citep{Vedaldi2010}.
+To do this, the patch is divided in 16 blocks, arranged in a 4x4 grid and centered on $T = (x,y)$, a location within the image.
+On each block, the bin consists of 45 degrees obtaining 8 aligning angles.
+
+The values of the image gradients aligned with each angle bin are accumulated through
+
+\begin{equation}
+ h(p,i,j) = 3 \cdot s \int w_\mathrm{ang}(\angle J(\mathbf{x}) - \theta_p)\, w_{ij}\left(\frac{\mathbf{x} - T}{3 s}\right)\, |J(\mathbf{x})|\, d\mathbf{x}
+\label{eq:histogram}
+\end{equation}
+
+\noindent where $s$ is the size of the local patch, $\angle J(\mathbf{x})$ is the gradient found at each of the 16 blocks of the patch aligned with the angle $\theta_p$.  
+This gives a descriptor of 128 dimension as shown on Figure \ref{fig:sampledescriptor}.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=16cm]{sampledescriptor.eps}
+\caption{The local patch is located around a sample point plotted on the temporary signal.
+All the sample points are interpolated using the Bresenham algorithm.
+The 4x4 grid can be seen on the descriptor and also arrows showing the dominant gradient on each block.}
+\label{fig:sampledescriptor}
+\end{figure}
 
@@
 
-Finally the classification is carried out by using the NBNN algorithm which can be described as follows.
+The classification is carried out by using the discriminative semi-supervised classification method, the Naive Bayes Nearest Neighbout (NBNN) \citep{Boiman2008}.
+% algorithm which can be described as follows.
 
-generative
+\begin{equation}
+\hat{C} = argmin_C \sum_{}^{} \left\lVert d_i - NN_C(d_i) \right\rVert ^2
+\end{equation}
 
-BCI Simulation vs k-fold cross validation.
+%BCI Simulation vs k-fold cross validation.
 It is quite known and understood that because the same experience will not generate the same markers in the EEG, BCI Simulation is sometimes better suited to understand how the device will operate on a real online application (the graeat variability of BCI).
@@
 \includegraphics[width=16cm]{subjectaveraged.eps}
-\caption{This is a figure, Schemes follow the same formatting.
-If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel.
-(\textbf{b}) Description of what is contained in the second panel.
-Figures should be placed in the main text near to the first time they are cited.
-A caption on a single line should be centered.}
+\caption{Point-To-Point grand averages of epochs obtained for hits (solid line) and no hits (dashed line) for each one of the 8 subjects for channel Cz.
+The P300 characteristic curve can be well identified on subjects 2 and 6.}
 \label{fig:subjectaveraged}
@@
 
- (CITE TAYLOR AND  FRANCOS PAPER).
+% (CITE TAYLOR AND  FRANCOS PAPER).
 
 \begin{table}[H]
-\caption{Results obtained for Accuracy levels by k-fold cross validation.
-The values reported for Cz are reproduced here for comparison.
-Cz values of BCI accuracies can be seen as long as the accuracy levels obtained for the best performing channels, for each subject.}
+\caption{Results obtained for Accuracy levels by 3-fold cross validation.
+The values reported by the dataset publishers for Cz are reproduced here for comparison.
+Additionally, channel Cz BCI accuracies can be seen as well as the performance levels obtained for the BPC, the best performing channel for each subject.}
 \centering
@@
 \toprule
-\textbf{Participants}	& \textbf{Original ACC at Cz}	& \textbf{ACC at Cz}	& \textbf{Best Performing Channel}	& \textbf{Performance}\\
+\textbf{Participants}	& \textbf{Original Cz}	& \textbf{ACC at Cz}	& \textbf{BPC}	& \textbf{Performance}\\
 \midrule
@@
 The ITR, or BTR, in the case of reactive BCIs \citep{WolpawJonathanR2012} strongly depends on the amount of signal averaging required to transmit a valid and robust selection.
-There is a trade-off that needs to be maintained between the amount of repetitions for each trial  \ref{fig:singletrial}
+There is a trade-off that needs to be balanced between the required number of repetitions for each trial to guarantee robust transmission and the the achieved speed of transmission affected by the repetions.
+As detailed in the table \ref{tab:singletrialreduction} this method is only reduced in a 20 percent in accuracy levels even by using only one sequence of repetitions.
 
-\begin{figure}[H]
+\begin{table}[H]
+\caption{Results obtained for Accuracy levels by performing a BCI simulation for the testing dataset.
+The BPC, best performing channel is shown and also performance levels for the 120, 60 and 12 repetitions.
+Finally the PR, performance reduction is described.
+Almost all the subjects performed around 70\% even with by averaging 2 signals for each class.}
 \centering
-\includegraphics[width=16cm]{singletriality.eps}
-\caption{This is a figure, Schemes follow the same formatting.
-If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel.
-(\textbf{b}) Description of what is contained in the second panel.
-Figures should be placed in the main text near to the first time they are cited.
-A caption on a single line should be centered.}
-\label{fig:singletrial}
-\end{figure}
+%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
+\begin{tabular}{cccccc}
+\toprule
+\textbf{Participants}	& \textbf{BPC}	& \textbf{ACC@120}	& \textbf{ACC@60}	&  \textbf{ACC@12} & \textbf{PR}\\
+\midrule
+1 & Cz &   0.80 &   0.82 &   0.72 & 9\% \\
+2 & PO7 &   0.93 &   0.78 &   0.75 & 18\%\\
+3 & Cz &   0.80 &   0.82 &   0.72 & 9\%\\
+4 & PO8 &   0.93 &   0.78 &   0.65 & 29\%\\
+5 & PO7 &   0.93 &   0.78 &   0.75 & 18\%\\
+6 & Cz &   0.80 &   0.82 &   0.72 & 9\%\\
+7 & PO7 &   0.93 &   0.78 &   0.75 & 18\%\\
+8 & PO7 &   0.93 &   0.78 &   0.75 & 18\%\\
+\bottomrule
+\end{tabular}
+\label{tab:singletrialreduction}
+\end{table}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-
-This method, different from other methods which is based on the nonlinearity of the gradient of histograms which can be used to detect
-
-is also based on how the image look like.
-
-SNR of p300 and how to detect it
-
-Check if you can use this to detect any kind of transient signal.
-
-Compare if it is possible with the descriptors from one subject, discriminate the others.
-
-Channal identification based on the metric distance between the bags
+%\section{Discussion}
+%
+%This method, different from other methods which is based on the nonlinearity of the gradient of histograms which can be used to detect 
+%
+%is also based on how the image look like.
+%
+%SNR of p300 and how to detect it
+%
+%Check if you can use this to detect any kind of transient signal.
+%
+%Compare if it is possible with the descriptors from one subject, discriminate the others.
+%
+%Channal identification based on the metric distance between the bags
 
@@
 
-With this result we showed a method to process EEG signals where the main characteristic can be both, rhytmic in nature as in motor imagery, and also transient and in time space, like the P300 event related potential.
+A method to characterize and classify EEG signals where the main characteristic can be both, rhythmic in nature as in motor imagery, and also transient in time space, like the P300 has been presented.
 
-If we avoid removing artefacts achieve a higher value of classification.
+Although single trial classification was not perfectly achieved, it has been shown that this method could be applied with a P300 Speller Matrix application with increased ITR.
 
-Si el desempeño da parecido hay que escribirlo ya que con menos valores se permite un ITR mayor.
+%Databases of descriptors.  Flexibility of the descriptor in terms of configuration and size (reference variants of SIFT on lower dimensiones).
 
-We also achieved single trial classification for just one subject.
+The adaptive behaviour of the algorithm make is well suited when the shape of the pattern elicited by the P300 response, does not conform to the predicted structure.
+This is because, the descriptors are directly based on how the signals actually looked like for the training and calibration step, and they do not require any prior knowledge about the signal.
+This is of particular relevance for studies on outliers populations as may be the case for people who suffered some neuro-degenerative disease like Lou Gehrig's disease.
 
-Databases of descriptors.  Flexibility of the descriptor in terms of configuration and size (reference variants of SIFT on lower dimensiones).
+The expanding of the understanding of this tool in order to automatically classify patterns in EEG, that are specifically identified by their shapes, is a prospect future work to be considered.
+It may also provide  assistance to physician or electroencephalographers to help them locate these EEG patterns particularly in long recording periods \citep{Hartman2005}, frequent in sleep research.
 
-Visual inspections enhancement tools.
-
-Understanding and expanding the use of this tool to automatically classify patterns in EEG, that are specifically identified by their shapes.
-This tool may provide the ability to help the physician to locate this patterns particularly in long-periods of time or long recordings.
-\citep{Hartman2005}
+Moreover, this method can be used as an alternate \textit{BCI predictor} \citep{Clerc} or as a tool for artefact removal (which is performed on many occasions by visually inspecting the signal).
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Subsection}
-
-\subsubsection{Subsubsection}
-
-Bulleted lists look like this:
-\begin{itemize}[leftmargin=*,labelsep=4mm]
-\item	First bullet
-\item	Second bullet
-\item	Third bullet
-\end{itemize}
-
-Numbered lists can be added as follows:
-\begin{enumerate}[leftmargin=*,labelsep=3mm]
-\item	First item
-\item	Second item
-\item	Third item
-\end{enumerate}
-
-The text continues here.
-
-\subsection{Figures, Tables and Schemes}
-
-All figures and tables should be cited in the main text as Figure 1, Table 1, etc.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=9cm]{subjectaveraged.eps}
-\caption{This is a figure, Schemes follow the same formatting.
-If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel.
-(\textbf{b}) Description of what is contained in the second panel.
-Figures should be placed in the main text near to the first time they are cited.
-A caption on a single line should be centered.}
-\end{figure}
-
-\subsection{Formatting of Mathematical Components}
-
-This is an example of an equation:
-
-\begin{equation}
-\mathbb{S}
-\end{equation}
-
-%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments).
-This ensures correct linenumbering.
-The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
-Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
-%% Example of a theorem:
-\begin{Theorem}
-Example text of a theorem.
-\end{Theorem}
-
-The text continues here. Proofs must be formatted as follows:
-
-%% Example of a proof:
-\begin{proof}[Proof of Theorem 1]
-Text of the proof. Note that the phrase `of Theorem 1' is optional if it is clear which theorem is being referred to.
-\end{proof}
-The text continues here.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-
-Authors should discuss the results and how they can be interpreted in perspective of previous studies and of the working hypotheses.
-The findings and their implications should be discussed in the broadest context possible.
-Future research directions may also be highlighted.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and Methods}
-
-Materials and Methods should be described with sufficient details to allow others to replicate and build on published results.
-Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers.
-Please disclose at the submission stage any restrictions on the availability of materials or information.
-New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.
-
-Research manuscripts reporting large datasets that are deposited in a publicly available database should specify where the data have been deposited and provide the relevant accession numbers.
-If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review.
-They must be provided prior to publication.
-
-Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-
-This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.
+%\subsection{Subsection}
+%
+%\subsubsection{Subsubsection}
+%
+%Bulleted lists look like this:
+%\begin{itemize}[leftmargin=*,labelsep=4mm]
+%\item	First bullet
+%\item	Second bullet
+%\item	Third bullet
+%\end{itemize}
+%
+%Numbered lists can be added as follows:
+%\begin{enumerate}[leftmargin=*,labelsep=3mm]
+%\item	First item
+%\item	Second item
+%\item	Third item
+%\end{enumerate}
+%
+%The text continues here.
 
@@
 %% optional
-\supplementary{The following are available online at www.mdpi.com/link, Figure S1: title, Table S1: title, Video S1: title.}
+%\supplementary{The following are available online at www.mdpi.com/link, Figure S1: title, Table S1: title, Video S1: title.}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{Juan Miguel Santos lead the project and conceived the idea. All the authors equally worked in the completion of this project.}
+\authorcontributions{This projects is part of a the first author's PhD Thesis which is directed by Juan Miguel Santos and codirected by Ana Julia Villar.}
 
@@
 
-\noindent EEG: Multidisciplinary Digital Publishing Institute\\
-BCI: Directory of open access journals\\
-SNR: Three letter acronym\\
-LD: linear dichroism}
+\noindent EEG: electroencephalography\\
+BCI: Brain Computer Interfaces\\
+SNR: Signal to Noise Ratio\\
+CNS: Central Nervous System\\
+ALS: Amyotrophic Lateral Sclerosis\\
+ERP: Event-Related Potential\\
+P300: Positive deflection of an Event-Related Potential which occurs 300 ms after onset of stimuli\\
+ITR: Information Transfer Rate}
 
@@
 %% optional
-\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
-\appendixsections{multiple} %Leave argument "multiple" if there are multiple sections.
+%\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
+%\appendixsections{multiple} %Leave argument "multiple" if there are multiple sections.
 Then a counter is printed ("Appendix A?).
@@
 and no counter is printed (?Appendix?).
-\appendix
-\section{}
-The appendix is an optional section that can contain details and data supplemental to the main text.
+%\appendix
+%\section{}
+%The appendix is an optional section that can contain details and data supplemental to the main text.
 For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data.
 Mathemtaical proofs of results not central to the paper can be added as an appendix.
-
-\section{}
-All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.
+%
+%\section{}
+%All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.
 
@@
 %% optional
-\sampleavailability{Samples of the compounds ...... are available from the authors.}
+%\sampleavailability{Samples of the compounds ...... are available from the authors.}
 

### 2016-12-20 coverletter.tex
Added ApeBrainStorm changes.

@@
 
-This submitted article encompass the work that we have been doing in a context of a PhD thesis and in the context of the setup of a Brain Computer Interfaces Laboratory at the Computer Engineering Department of the ITBA University in Buenos Aires Argentina.
+en el contexto de nuestra labor en el bci lab en el departamento bla bla b
+
+This submitted article encompass the work that we have been doing and in the context of the setup of a Brain Computer Interfaces Laboratory at the Computer Engineering Department of the ITBA University in Buenos Aires Argentina.
+
+Son areas de investigacion en las cuales nosotros podemos llegar a contribuir como ingenieros e investigadores a los efectos de brindar nuevos metodos y herramientas que favorezcan la comunicación o el tratamiento de individuos que han sido dañados o afectados por enfermedades neuro degenerativas
 

### 2016-12-20 article.tex
Final Version before deadline

@@
 % Full title of the paper (Capitalized)
-\Title{EEG characterization and classification based on histograms of visual gradients}
+\Title{EEG characterization and classification based on histograms of image gradients}
 
@@
 
-Although impressive advances in neuroimagining techniques (particularly radio-nuclear and radiological scanning methods) \citep{Schomer2010} have diminished the prospects of the traditional Electroencephalography (EEG), the advent and development of digitalized devices has pressed for a revamping of this hundredth years old technology.
+Although recent advances in neuroimagining techniques (particularly radio-nuclear and radiological scanning methods) \citep{Schomer2010} have diminished the prospects of the traditional Electroencephalography (EEG), the advent and development of digitalized devices has pressed for a revamping of this hundredth years old technology.
 Their versatility, ease of use, temporal resolution, ease of development and fabrication, and its proliferation as consumer devices, are pushing EEG to become the de-facto non invasive method to access and harness brain information.
@@
 
+Besides novel applications of BCI, the goal of the entire discipline is to provide communication assistance to people affected by neuro-degenerative diseases.
+It is understood that for people suffering from ALS, even though this pathology do not harm the CNS, their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns alterations \citep{Nijboer2009,Riener2014}.
+
+BCI has gained mainstream public awareness and recently even held a peculiar Olympics \citep{Riener2014} and been broadcasted during the inauguration ceremony of the 2014 Soccer World Cup.
+New developments have overcome the out-of-the-lab high-bar and they are starting to be used in real world environments \citep{Huggins2016}.
+However, they still lack the necessary robustness, and its performance is well behind any other method of human computer interaction, including any kind of detection of remaining muscular movement \citep{Clerc}.
+
 In this work a new method to characterize EEG signals is presented, expanded and detailed.
 Its validity is verified by processing off-line data for ALS patients.
-This is the continuation of the work previously presented in \citep{Ramele2016}, where it was applied to rhythmic patterns, and that it can be extended to describe transient events like those produced by the P300\citep{Knuth2006}.
+This is the continuation of the work previously presented in \citep{Ramele2016}, where it was applied to rhythmic patterns, and that it can be extended to describe transient events like those produced by the P300 \citep{Knuth2006}.
 
-The method is based on the morphological analysis of the shape of the EEG signal\citep{Alvarado-Gonzalez2016,Yamaguchi2009} and was inspired by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting  raw signals.
+The method is based on the morphological analysis of the shape of the EEG signal \citep{Alvarado-Gonzalez2016,Yamaguchi2009} and was inspired by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting  raw signals \citep{Hartman2005}.
 
@@
 
-In brief, this paper reports a straightforward method to, first characterize EEG signals based on the identification of their morphological shapes, and how this characterization can be used to implement a BCI scheme to identify Event Related Potentials, particularly the well-known P300, on an off-line and public dataset.
+In brief, this paper reports a straightforward method to, first characterize EEG signals based on the identification of their morphological shapes, and how this characterization can be used to implement a BCI scheme to identify Event Related Potentials of ALS patients, particularly the well-known P300, on an off-line and public dataset.
 
 \section{Materials and Methods}
-
-Besides new novel applications of BCI, the goal of the entire discipline is to provide communication assistance to people affected by neuro-degenerative diseases.
-It is understood that for people suffering from ALS, even though this pathology do not harm the CNS, their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns \citep{Nijboer2009,Riener2014}.
-
-BCI has gained mainstream public awareness and recently even held a peculiar Olympics \citep{Riener2014} and been broadcasted during the inauguration of the last Soccer World Cup.
-New developments have overcome the out-of-the-lab high-bar and they are starting to be used in real world environments\citep{Huggins2016}.
-However, they still lack the necessary robustness, and its performance is well behind any other method of human computer interaction, including any kind of detection of remaining muscular movement \citep{Clerc}.
 
@@
 A trial, for this experiment, was defined as every attempt to select a letter from the speller, and it was composed of 10 repetitions of flashes of 6 rows and 6 columns of the traditional 6x6 P300 matrix , yielding 120 repetitions.
-Flashing of row/columns was performed for 0.125 s, following by a resting period of the same size.
-After 120 repetions and inter-trial pause was included before resuming with the following letter.
+Flashing of row/columns was performed for 0.125 s, following by a resting period of the same length.
+After 120 repetitions an inter-trial pause was included before resuming with the following letter.
 
-The gathered dataset was sampled at 256 Hz and consisted of EEG matrix for electrode channels Fz,Cz,Pz,Oz,P3,P4,PO7 and PO8, identified according to the 10-20 International System,  for each one of the 8 subjects.
+The gathered dataset was sampled at 256 Hz and consisted of an EEG matrix for electrode channels Fz,Cz,Pz,Oz,P3,P4,PO7 and PO8, identified according to the 10-20 International System,  for each one of the 8 subjects.
 
@@
 
+The proposed algorithm was developed to perform a binary classification task, determining the appearance on each corresponding signal segment, of the P300 response.
+If it was present, the segment was labelled as class 1, a \textit{hit}, and its absent was identified as class 2, a \textit{no-hit}.
+
 \subsubsection{Preprocessing}
@@
 The mandatory first step is to enhance the SNR of the P300 pattern above the level of basal EEG.
-The processing pipeline starts with applying a notch filter to the raw signal, a decimation with a FIR filter of order 30 from the original 256 Hz to 32 Hz and finally applying a 4th degree 10 Hz lowpass Butterworth filter.
+The processing pipeline starts by applying a notch filter to the raw signal, a decimation with an FIR filter of order 30 from the original 256 Hz to 32 Hz and finally applying a 4th degree 10 Hz lowpass Butterworth filter.
 
@@
 
-From every segment of 12 flashings, two epoch corresponding to a hit and two matching a no-hit are constructed as segments of 1s length (256 sample points).
+From every segment of 12 flashings, two epoch were obtained when the signal was labelled as \textit{hit} and two when the signal was labelled as a \textit{no-hit} of 1s length both (256 sample points).
 
-These segments, 20 for each class, are later point-to-point averaged for the whole 120 repetitions group.
-This finally lead to 2 averaged signals for each of the 5 letters, for the 7 words, comprising a set of 70 averaged signals\citep{Liang2008}.
+These segments, 20 for each class, are later point-to-point averaged for the whole trial.
+This finally lead to 2 averaged signals for each of the 5 letters, for the 7 words, comprising a set of 70 averaged signals \citep{Liang2008}.
 
@@
 
-The central part of this method is the feature generation: the histogram of visual gradients.  To do so, the first step is the transformation of the signal into a temporary binary image.
+The central part of this method is the feature generation: the histogram of image gradients.  To do so, the first step is the transformation of the signal into a temporary binary image.
 
@@
 
-Finally, the histogram of gradients, which is based on a technique from Computer Vision, the SIFT \citep{Lowe2004} descriptor, is calculated on local image patches centered on sample points \citep{Vedaldi2010}.
+Finally, the histogram of gradients, which is based on the SIFT \citep{Lowe2004} descriptor, a technique from Computer Vision, is calculated on local image patches centered on sample points \citep{Vedaldi2010}.
 To do this, the patch is divided in 16 blocks, arranged in a 4x4 grid and centered on $T = (x,y)$, a location within the image.
-On each block, the bin consists of 45 degrees obtaining 8 aligning angles.
-
-The values of the image gradients aligned with each angle bin are accumulated through
+The values of the image gradients for each angle bin are accumulated through
 
@@
 
-\noindent where $s$ is the size of the local patch, $\angle J(\mathbf{x})$ is the gradient found at each of the 16 blocks of the patch aligned with the angle $\theta_p$.  
-This gives a descriptor of 128 dimension as shown on Figure \ref{fig:sampledescriptor}.
+\noindent where $s$ is the size of the local patch, which can be converted into pixels by doing $ Px = 4 \cdot 3 \cdot s $ , $\angle J(\mathbf{x})$ is the angle of the gradient vector $ J(\mathbf{x}) $  found at each one of the 16 blocks of the patch, whereas  $\theta_p$ is one of the bin angles.
+On the other hand $ x = (x_i, x_j) $ corresponds to each one of the pixels of the local patch, and $ w_\mathrm{ang}  $  and $ w_{ij} $ are linear interpolation functions \citep{Lowe2004}.
+
+Summarizing, each block of the patch defines 8 different bins of the histogram.
+On each one of these bins  the magnitude of the image gradients at different angles is summed up (from 0 to 360, on steps of 45 degrees).
+As the patch is 4x4, it gives a descriptor of 128 dimension as shown on Figure \ref{fig:sampledescriptor}.
 
@@
 \centering
-\includegraphics[width=16cm]{sampledescriptor.eps}
-\caption{The local patch is located around a sample point plotted on the temporary signal.
+\includegraphics[width=6cm]{sampledescriptor.png}
+\caption{The local patch is located around a sample point plotted on the temporary image.
 All the sample points are interpolated using the Bresenham algorithm.
-The 4x4 grid can be seen on the descriptor and also arrows showing the dominant gradient on each block.}
+The 4x4 block grid can be seen on the patch and also arrows showing the dominant gradient on each block.}
 \label{fig:sampledescriptor}
@@
 
-The classification is carried out by using the discriminative semi-supervised classification method, the Naive Bayes Nearest Neighbout (NBNN) \citep{Boiman2008}.
-% algorithm which can be described as follows.
-
-\begin{equation}
-\hat{C} = argmin_C \sum_{}^{} \left\lVert d_i - NN_C(d_i) \right\rVert ^2
-\end{equation}
+The classification is carried out by using a discriminative semi-supervised classification method, the Naive Bayes Nearest Neighbout (NBNN) \citep{Boiman2008} and the first step consist in extracting from the training set labelled descriptor for each class which are thus grouped together in two KD-tree \citep{Vedaldi2010} database structures.
 
@@
 
+While doing classification based on local features, which encode partial information, one problem that frequently arises is how to ensemble the individual classification of each local characteristic to the image that were used to derive those descriptors.
+
+The NBNN algorithm tackles this problem by comparing each image against a whole classification class which is characterized by the set of descriptors that are closest to each one of the descriptors obtained for the query image.
+The algorithm is based on \ref{eq:classification},
+
+\begin{equation}
+\hat{C} = argmin_C \sum_{}^{} \left\lVert d_i - NN_C(d_i) \right\rVert ^2
+\label{eq:classification}
+\end{equation}
+
+\noindent where the predicted class $\hat{C}$ of a query image is calculated as the class $ C $ that minimize the summation of the $ L^2 $ distance between each descriptor $ d_i $ that belongs to the query image and its corresponding near neighbour $ NN_C(d_i) $ descriptor from each class.
+
+\subsubsection{Parameters}
+
+This method's parameters were selected according to the experimental protocol.
+As the P300 event latency varies greatly between subjects, it is necessary to provide a descriptor that will be able to capture an entire transient event.
+Equations \ref{eq:mapping1} and \ref{eq:mapping2} can be used to map the original signal parameters to local image patch structure.
+
+\begin{equation}
+s = \frac{\Delta \mu V}{4 \cdot 3} \cdot \gamma 
+\label{eq:mapping1}
+\end{equation}
+
+\begin{equation}
+s = \frac{\lambda \cdot Fs}{4 \cdot 3} \cdot \gamma
+\label{eq:mapping2}
+\end{equation}
+
+\noindent where $ Fs $ is the sample frequency of the EEG signal, $ \lambda $ is the length in seconds covered by the patch, and $ \mu V $ correspond to the amplitude in microvolts that can be covered by the height of the patch.
+By using a $ s = 3 $ and a double scale of the image $ \gamma = 2 $ this gives the local patch, and the descriptor, the ability to identify events of 18 $ \mu V $ of amplitude of $ \lambda = 0.5 s $.
+Finally, descriptors locations $ T $ were selected as suggested by the dataset publisher \citep{Riccio2013} between 0.2s and 0.7s after the onset of the stimuli.
+
+%\begin{figure}[H]
+%\centering
+%\includegraphics[width=16cm]{localpatch.png}
+%\caption{The local patch is mapped to the original image by their correspondence .}
+%\label{fig:sampledescriptor}
+%\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-In Figure \ref{fig:subjectaveraged} the grand average (point-to-point) for all the subjects is shown.
+In Figure \ref{fig:subjectaveraged} the grand average (point-to-point) for all the subjects using the whole trial (i.e.
+120 repetitions) is shown.
 The P300 characteristic curve can be seen particularly in subjects 2, and 6 and in a lesser extend in the remaining subjects.
-In order to obtain a valid binary classification on averaged signals, particular care was observed to avoid unbalanced samples, because that may introduce a bias in the classification procedure (signal variance is proportional to the square root of the number of samples and the procedure would be discriminating signals with different variances).Particularly in the case of the P300 response, the oddball paradigm requires that one of the stimulus need to be infrequent so that will unavoidably force that the data will be unbalanced (that is, an unequal number of trials in each condition).
+In order to obtain a valid binary classification on averaged signals, particular care was observed to avoid unbalanced samples, because that may introduce a bias in the classification procedure (the variance of average signals is inversely proportional to the number of samples and the procedure would be discriminating signals with different variances).Particularly in the case of the P300 response, the oddball paradigm requires that one of the stimulus need to be infrequent so that will unavoidably force that the data will be unbalanced (that is, an unequal number of trials in each condition).
 
@@
 \caption{Point-To-Point grand averages of epochs obtained for hits (solid line) and no hits (dashed line) for each one of the 8 subjects for channel Cz.
-The P300 characteristic curve can be well identified on subjects 2 and 6.}
+The P300 characteristic curve can be well identified particularly on subjects 2 and 6.}
 \label{fig:subjectaveraged}
@@
 Results are shown in Table \ref{tab:results} where BCI accuracy
-for the Cz channel is calculated and compared with the values obtained by the published of the dataset.
+for the Cz channel is calculated and compared with the values obtained by the published of the dataset  \citep{Riccio2013}.
 Additionally the best performing channel is informed and its accuracy value.
@@
 The values reported by the dataset publishers for Cz are reproduced here for comparison.
-Additionally, channel Cz BCI accuracies can be seen as well as the performance levels obtained for the BPC, the best performing channel for each subject.}
+Additionally, channel Cz BCI accuracies can be seen as well as the performance levels and their standard deviation obtained for the BPC, the best performing channel for each subject.}
 \centering
@@
 \toprule
-\textbf{Participants}	& \textbf{Original Cz}	& \textbf{ACC at Cz}	& \textbf{BPC}	& \textbf{Performance}\\
+\textbf{Participant}	& \textbf{Original Cz}	& \textbf{ACC at Cz}	& \textbf{BPC}	& \textbf{Performance}\\
 \midrule
@@
 The BPC, best performing channel is shown and also performance levels for the 120, 60 and 12 repetitions.
-Finally the PR, performance reduction is described.
-Almost all the subjects performed around 70\% even with by averaging 2 signals for each class.}
+Finally the PR, performance reduction as a rate of reduction compared to the accuracy obtained using the entire trial, is described.
+Almost all the subjects performed around 70\% even by averaging only 2 epochs for each class.}
 \centering
@@
 \toprule
-\textbf{Participants}	& \textbf{BPC}	& \textbf{ACC@120}	& \textbf{ACC@60}	&  \textbf{ACC@12} & \textbf{PR}\\
+\textbf{Participant}	& \textbf{BPC}	& \textbf{ACC@120}	& \textbf{ACC@60}	&  \textbf{ACC@12} & \textbf{PR}\\
 \midrule
@@
 P300: Positive deflection of an Event-Related Potential which occurs 300 ms after onset of stimuli\\
-ITR: Information Transfer Rate}
+ITR: Information Transfer Rate\\
+BTR: Bit Transfer Rate}
 

### 2016-12-22 article.tex
Subo correcciones de prosa que habían quedado mal y agregué un pequeño detalle de que es P300 que nos había faltado

@@
 
-One very noteworthy aspect of this novel communication channel is the ability to volitionally transmit information from the Central Nervous System (CNS) to a computer device and from there use that information to control a wheelchair \citep{Carlson2013}, as input to a speller application \citep{Guger2009a}, or as aiding tool in a rehabilitation procedure \citep{Jure2016}.
+One noteworthy aspect of this novel communication channel is the ability to volitionally transmit information from the Central Nervous System (CNS) to a computer device and from there use that information to control a wheelchair \citep{Carlson2013}, as input to a speller application \citep{Guger2009a}, or as aiding tool in a rehabilitation procedure \citep{Jure2016}.
 The holly grail of BCI is to implement a new complete and alternative pathway to restore lost locomotion \citep{WolpawJonathanR2012}.
@@
 EEG signals are remarkably complex and have been characterized as a multichannel non-stationary stochastic process and even considered random enough as to be used as a source of pseudo random number generator \citep{Chen2014}.
-Additionally, they have a high variability between different subjects and even between different moments for the same subject, requiring an adaptive and co-adaptive calibration and learning procedures \citep{Clerc}.
+Additionally, they have high variability between different subjects and even between different moments for the same subject, requiring an adaptive and co-adaptive calibration and learning procedures \citep{Clerc}.
 Hence, this imposes an outstanding challenge that is necessary to overcome in order to extract information from raw EEG signals.
 
-EEG markers \citep{Clerc} that can be used to transmit volitionally information are limited, and each one of them has a particular set or combination of appropriate methods to decode them.
+EEG markers \citep{Clerc} that can be used to volitionally transmit information are limited, and each one of them has a particular set or combination of appropriate methods to decode them.
 Inevitably, it is often necessary to implement many distinct and specialized algorithmic methods, to filter the signal, enhance its SNR, and try to determine some meaning out of it.
@@
 
-%This paper reports on a study done to address
-%the hypothesis that communication task complexity
-%(copy-spelling vs. novel text generation of a picture
-%description) is correlated with BCI performance accuracy
-%and P300 characteristics.
-
 In brief, this paper reports a straightforward method to, first characterize EEG signals based on the identification of their morphological shapes, and how this characterization can be used to implement a BCI scheme to identify Event Related Potentials of ALS patients, particularly the well-known P300, on an off-line and public dataset.
@@
 The protocol is explained in \citep{Riccio2013} but can be summarized as follows:  8 subjects with confirmed diagnoses but on different stages of ALS disease, were recruited and accepted to perform the experiments.
-The P300 task objective was meant to spell 7 words of 5 letter each.
+The P300 \citep{Knuth2006} is a positive deflection of the EEG signal which occurs 300 ms after the onset of a rare and deviant stimulus that the subject is expected to attend.
+It is produced under the oddball paradigm \cite{WolpawJonathanR2012} and tough it is quite consistent across different subjects, it is small ( $ \pm 5 \mu V $ ) compared to the ongoing EEG activity.
+The P300 detection task designed for this experiment was meant to spell 7 words of 5 letter each, using the traditional P300 Speller Matrix, where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.
 The first 3 runs were used for training and the remaining 4 for testing with visual feedback.
-A trial, for this experiment, was defined as every attempt to select a letter from the speller, and it was composed of 10 repetitions of flashes of 6 rows and 6 columns of the traditional 6x6 P300 matrix , yielding 120 repetitions.
+A trial, for this experiment, was defined as every attempt to select a letter from the speller, and it was composed of signal segments corresponding to 10 repetitions of flashes of 6 rows and 6 columns of the traditional 6x6 P300 matrix , yielding 120 repetitions.
 Flashing of row/columns was performed for 0.125 s, following by a resting period of the same length.
@@
 
-The proposed algorithm was developed to perform a binary classification task, determining the appearance on each corresponding signal segment, of the P300 response.
+The proposed algorithm was developed to perform a binary classification task, determining the appearance on each corresponding signal segment of the P300 response.
 If it was present, the segment was labelled as class 1, a \textit{hit}, and its absent was identified as class 2, a \textit{no-hit}.
@@
 For each word, each letter, a segment of EEG signal is extracted that encompass 10 repetitions of 12 flashings, 6 rows and 6 columns.
-For every 12 stimuli flashing, i.e.
+For every 12 flashing stimuli, i.e.
 one complete sequence of repetitions of each row/column, a basic artefact removal procedure of removing the entire segment if any signal goes above 70 $\mu$V is applied.
@@
 These segments, 20 for each class, are later point-to-point averaged for the whole trial.
-This finally lead to 2 averaged signals for each of the 5 letters, for the 7 words, comprising a set of 70 averaged signals \citep{Liang2008}.
+It is important to remark that only 2 out of 10 available segments marked as a \textit{no-hit} were extracted in order to avoid generating unbalanced averaging signals (see Results section \ref{section:results}).
+This finally led to 2 averaged signals for each of the 5 letters, for the 7 words, comprising a set of 70 averaged signals \citep{Liang2008}.
 
@@
 All the sample points are interpolated using the Bresenham algorithm.
-The 4x4 block grid can be seen on the patch and also arrows showing the dominant gradient on each block.}
+The 4x4 block grid can be seen on the patch as well as the green arrows showing the dominant gradient on each block.}
 \label{fig:sampledescriptor}
@@
 
-The classification is carried out by using a discriminative semi-supervised classification method, the Naive Bayes Nearest Neighbout (NBNN) \citep{Boiman2008} and the first step consist in extracting from the training set labelled descriptor for each class which are thus grouped together in two KD-tree \citep{Vedaldi2010} database structures.
+The classification is carried out by using a discriminative semi-supervised classification method, the Naive Bayes Nearest Neighbout (NBNN) \citep{Boiman2008} and the first step consists in extracting labelled descriptors from the training set for each class which are thus grouped together in two KD-tree \citep{Vedaldi2010} database structures.
 
@@
 
-\noindent where $ Fs $ is the sample frequency of the EEG signal, $ \lambda $ is the length in seconds covered by the patch, and $ \mu V $ correspond to the amplitude in microvolts that can be covered by the height of the patch.
-By using a $ s = 3 $ and a double scale of the image $ \gamma = 2 $ this gives the local patch, and the descriptor, the ability to identify events of 18 $ \mu V $ of amplitude of $ \lambda = 0.5 s $.
-Finally, descriptors locations $ T $ were selected as suggested by the dataset publisher \citep{Riccio2013} between 0.2s and 0.7s after the onset of the stimuli.
+\noindent where $ Fs $ is the sample frequency of the EEG signal, $ \lambda $ is the length in seconds covered by the patch, and $ \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch.
+By using $ s = 3 $ and a double scale of the image $ \gamma = 2 $ this gives the local patch, and the descriptor, the ability to identify events of 18 $ \mu V $ of amplitude of $ \lambda = 0.5 s $.
+Finally, descriptors locations $ T $ were selected as suggested by the dataset publisher \citep{Riccio2013} between 0.2s and 0.7s after the onset of the stimulus.
 
@@
 \section{Results}
-
+\label{section:results}
 In Figure \ref{fig:subjectaveraged} the grand average (point-to-point) for all the subjects using the whole trial (i.e.
@@
 The P300 characteristic curve can be seen particularly in subjects 2, and 6 and in a lesser extend in the remaining subjects.
-In order to obtain a valid binary classification on averaged signals, particular care was observed to avoid unbalanced samples, because that may introduce a bias in the classification procedure (the variance of average signals is inversely proportional to the number of samples and the procedure would be discriminating signals with different variances).Particularly in the case of the P300 response, the oddball paradigm requires that one of the stimulus need to be infrequent so that will unavoidably force that the data will be unbalanced (that is, an unequal number of trials in each condition).
+In order to obtain a valid binary classification on averaged signals, particular care was observed to avoid unbalanced samples, because that may introduce a bias in the classification procedure (the variance of average signals is inversely proportional to the number of samples and the procedure would be discriminating signals with different variances).
+Especially in the case of the P300 response, the oddball paradigm requires that one of the stimuli need to be infrequent so that will unavoidably force the data to be unbalanced (i.e.
+an unequal number of trials in each condition) \citep{Tibon2015}.
 
@@
 Results are shown in Table \ref{tab:results} where BCI accuracy
-for the Cz channel is calculated and compared with the values obtained by the published of the dataset  \citep{Riccio2013}.
-Additionally the best performing channel is informed and its accuracy value.
+for the Cz channel is calculated and compared with the values obtained by the publisher of the dataset  \citep{Riccio2013}.
+Additionally the best performing channel is informed as well as its accuracy value.
 It is of particular interest that using this method, the best performing channel was not always Cz, and instead occipital channels PO8 and PO7 showed very good performance indeed \citep{Huggins2016,Jure2016}.
 
-% (CITE TAYLOR AND  FRANCOS PAPER).
-
 \begin{table}[H]
-\caption{Results obtained for Accuracy levels by 3-fold cross validation.
+\caption{Accuracy levels obtained by a 3-fold cross validation.
 The values reported by the dataset publishers for Cz are reproduced here for comparison.
-Additionally, channel Cz BCI accuracies can be seen as well as the performance levels and their standard deviation obtained for the BPC, the best performing channel for each subject.}
+Additionally, BCI accuracies for channel Cz can be seen as well as performance levels and their standard deviation obtained for the BPC, the best performing channel for each subject.}
 \centering
@@
 The ITR, or BTR, in the case of reactive BCIs \citep{WolpawJonathanR2012} strongly depends on the amount of signal averaging required to transmit a valid and robust selection.
-There is a trade-off that needs to be balanced between the required number of repetitions for each trial to guarantee robust transmission and the the achieved speed of transmission affected by the repetions.
-As detailed in the table \ref{tab:singletrialreduction} this method is only reduced in a 20 percent in accuracy levels even by using only one sequence of repetitions.
+There is a trade-off that needs to be balanced between the required number of repetitions for each trial to guarantee robust transmission and the the achieved speed of transmission affected by the repetitions.
+As detailed in the table \ref{tab:singletrialreduction} by applying this method, accuracy levels are only reduced in 20 percent even by using only one sequence of 12 repetitions (i.e.
+an entire flashing of the whole matrix).
 
 \begin{table}[H]
-\caption{Results obtained for Accuracy levels by performing a BCI simulation for the testing dataset.
+\caption{Accuracy levels obtained by performing a BCI simulation on the testing dataset (4 letters).
 The BPC, best performing channel is shown and also performance levels for the 120, 60 and 12 repetitions.
@@
 
-The adaptive behaviour of the algorithm make is well suited when the shape of the pattern elicited by the P300 response, does not conform to the predicted structure.
-This is because, the descriptors are directly based on how the signals actually looked like for the training and calibration step, and they do not require any prior knowledge about the signal.
-This is of particular relevance for studies on outliers populations as may be the case for people who suffered some neuro-degenerative disease like Lou Gehrig's disease.
+The adaptive behaviour of the algorithm make it well suited when the shape of the pattern elicited by the P300 response, does not conform to the predicted structure.
+This is due to the fact that the descriptors are directly based on how the signals actually looked like for the training and calibration step, and they do not require any prior knowledge about the signal.
+This is of particular relevance for studies on outliers populations as may be the case for people who suffered some form of neuro-degenerative disorder like Lou Gehrig's disease.
 
@@
 ERP: Event-Related Potential\\
-P300: Positive deflection of an Event-Related Potential which occurs 300 ms after onset of stimuli\\
+P300: Positive deflection of an Event-Related Potential which occurs 300 ms after onset of stimulus\\
 ITR: Information Transfer Rate\\
-BTR: Bit Transfer Rate}
+BTR: Bit Transfer Rate\\
+SIFT: Scale Invariant Feature Transform\\
+HOG: Histogram Of Gradients}
 

### 2016-12-22 article.tex
Added explanation of P300 phenomena and some minor typos

@@
 
-The proposed algorithm was developed to perform a binary classification task, determining the appearance on each corresponding signal segment of the P300 response.
+The proposed algorithm was developed to perform a binary classification task, determining the appearance or not of the P300 response on each corresponding signal segment.
 If it was present, the segment was labelled as class 1, a \textit{hit}, and its absent was identified as class 2, a \textit{no-hit}.

### 2017-02-18 article.tex
Corrections

@@
 % Abstract (Do not use inserted blank lines, i.e. \\) 
-\abstract{The analysis of Electroencephalograpy (EEG) signals for Amyotrophic Lateral Sclerosis (ALS) patients is of ulterior importance to elucidate different patterns or markers that could be potentially used to aid in diagnosis or treatment.
-Additionally, it will improve the implementation of Brain Computer Interface (BCI) systems which may provide alternative pathways to transmit volitional information to enhance patient's quality of life.
-Of particular interests are those which are based on the recognition of Event-Related Potentials (ERP) which are shown to be more robust to the unavoidable physical impairment which comes with this motor neuron disease.
+\abstract{The analysis of Electroencephalograpy (EEG) signals for Amyotrophic Lateral Sclerosis (ALS) patients is of ulterior importance to elucidate different patterns or markers that could improve the implementation of Brain Computer Interface (BCI).
+These systems are meant to provide alternative pathways to transmit volitional information which could potentially enhance patient's quality of life.
+Of particular interests are those which are based on the recognition of Event-Related Potentials (ERP), shown to be more robust to the unavoidable physical impairment which comes with this motor neuron disease.
 This work mimics what electroencephalographers have been doing clinically, visually inspecting and categorizing phenomena within the EEG, and provides a framework to analyse and characterize EEG signals, with a focus on the P300, an ERP elicited by the oddball paradigm of rare events.
@@
 Although recent advances in neuroimagining techniques (particularly radio-nuclear and radiological scanning methods) \citep{Schomer2010} have diminished the prospects of the traditional Electroencephalography (EEG), the advent and development of digitalized devices has pressed for a revamping of this hundredth years old technology.
-Their versatility, ease of use, temporal resolution, ease of development and fabrication, and its proliferation as consumer devices, are pushing EEG to become the de-facto non invasive method to access and harness brain information.
+Their versatility, ease of use, temporal resolution, ease of development and fabrication, and its proliferation as consumer devices, are pushing EEG to become the de-facto non invasive portable or ambulatory method to access and harness brain information \cite{DeVos2014}
 
@@
 Inevitably, it is often necessary to implement many distinct and specialized algorithmic methods, to filter the signal, enhance its SNR, and try to determine some meaning out of it.
+NEED REVIEW Patients suffering from ALS their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns alterations \citep{Nijboer2009,Riener2014}.
 
-Besides novel applications of BCI, the goal of the entire discipline is to provide communication assistance to people affected by neuro-degenerative diseases.
+%Besides novel applications of BCI, the goal of the entire discipline is to provide communication assistance to people affected by neuro-degenerative diseases.
 It is understood that for people suffering from ALS, even though this pathology do not harm the CNS, their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns alterations \citep{Nijboer2009,Riener2014}.
@@
 
-To verify the validity of the proposed framework and method, the public dataset 008-2014  \citep{Riccio2013} published on the BNCI-Horizon website \citep{Brunner2014} was used to perform a binary classification task on the provided signals.
+To verify the validity of the proposed framework and method, the public dataset 008-2014  \citep{Riccio2013} published on the BNCI-Horizon website \citep{Brunner2014} by  IRCCS Fondazione Santa Lucia, was used to perform a binary classification task on the provided signals.
 The algorithm was implemented using the VLFeat  \citep{Vedaldi2010} Computer Vision libraries on MATLAB 2014a (Mathworks Inc., Natick, MA, USA).
@@
 The P300 \citep{Knuth2006} is a positive deflection of the EEG signal which occurs 300 ms after the onset of a rare and deviant stimulus that the subject is expected to attend.
-It is produced under the oddball paradigm \cite{WolpawJonathanR2012} and tough it is quite consistent across different subjects, it is small ( $ \pm 5 \mu V $ ) compared to the ongoing EEG activity.
+It is produced under the oddball paradigm \cite{WolpawJonathanR2012} and though it is quite consistent across different subjects, it is small ( $ \pm 5 \mu V $ ) compared to the ongoing EEG activity.
 The P300 detection task designed for this experiment was meant to spell 7 words of 5 letter each, using the traditional P300 Speller Matrix, where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.
@@
 
-The gathered dataset was sampled at 256 Hz and consisted of an EEG matrix for electrode channels Fz,Cz,Pz,Oz,P3,P4,PO7 and PO8, identified according to the 10-20 International System,  for each one of the 8 subjects.
+The recorded dataset was sampled at 256 Hz and consisted of an EEG matrix for electrode channels Fz,Cz,Pz,Oz,P3,P4,PO7 and PO8, identified according to the 10-20 International System,  for each one of the 8 subjects.
 
@@
 The mandatory first step is to enhance the SNR of the P300 pattern above the level of basal EEG.
-The processing pipeline starts by applying a notch filter to the raw signal, a decimation with an FIR filter of order 30 from the original 256 Hz to 32 Hz and finally applying a 4th degree 10 Hz lowpass Butterworth filter.
+The processing pipeline starts by applying a notch filter to the raw signal, a 4th degree 10 Hz lowpass Butterworth filter and finally a decimation with an FIR filter of order 30 from the original 256 Hz to 32 Hz.
 
@@
 
-These segments, 20 for each class, are later point-to-point averaged for the whole trial.
+These segments, 20 for each class, are later point-to-point averaged for the whole 120 repetitions.
 It is important to remark that only 2 out of 10 available segments marked as a \textit{no-hit} were extracted in order to avoid generating unbalanced averaging signals (see Results section \ref{section:results}).
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Discussion}
-%
-%This method, different from other methods which is based on the nonlinearity of the gradient of histograms which can be used to detect 
-%
-%is also based on how the image look like.
-%
-%SNR of p300 and how to detect it
-%
-%Check if you can use this to detect any kind of transient signal.
-%
-%Compare if it is possible with the descriptors from one subject, discriminate the others.
-%
-%Channal identification based on the metric distance between the bags
+\section{Discussion}
+
+This method, different from other methods which is based on the nonlinearity of the gradient of histograms which can be used to detect
+
+is also based on how the image look like.
+
+SNR of p300 and how to detect it
+
+Check if you can use this to detect any kind of transient signal.
+
+Compare if it is possible with the descriptors from one subject, discriminate the others.
+
+Channal identification based on the metric distance between the bags
 
